### 2016-01-07 reports/paper-hdrda.Rnw
Some edits from Phil.
DMY is going to have a look at this version now.

\documentclass{article}

\usepackage{graphicx, amsmath, amssymb, bm, mathtools, amsthm, natbib, setspace, algorithm2e}

\bibpunct{(}{)}{;}{a}{,}{,}

\newcommand{\xbar}{\bar{\bm x}}
\newcommand{\tr}{\text{tr}}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\newtheorem{cor}{Corollary}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{thm}{Theorem}

\pagestyle{plain}
%----------------Page dimensions ----------------
\oddsidemargin 0.0in
\evensidemargin 0.0in
\topmargin -0.75in
\leftmargin 0in
\headheight 0.0in
\headsep 0.5in
\footskip 0.5in
\footnotesep 0.0in
\textwidth 6.5in
\textheight 9.5in

\begin{document}

\title{High-Dimensional Regularized Discriminant Analysis}

\author{John A. Ramey, Caleb K. Stein, Phil D. Young, and Dean M. Young}

\maketitle

\doublespacing

\begin{abstract}

\citeauthor{Friedman:1989tm} proposed the popular regularized discriminant
analysis (RDA) classifier that utilizes a biased covariance-matrix estimator to
partially pool the sample covariance matrices from linear and quadratic
disriminant analysis before shrinking the resulting estimator towards a scaled
identity matrix. The RDA classifier's two tuning parameters are typically
estimated via a computationally burdensome cross-validation procedure from a
grid of candidate values. We reformulate the RDA classifier in the small-sample,
high-dimensional setting and then show that the classification decision rule is
equivalent to a classifier in a subspace having a much lower dimension. As a
result, the dimension reduction involved yields a substantial reduction in
computation during model selection. Our parameterization also offers
interpretability previously lacking with the RDA classifier. We demonstrate with
three artificial and four real high-dimensional data sets that our proposed
classifier is often superior to several recently proposed sparse and regularized
classifiers in terms of classification accuracy. Finally, we provide an
implementation of our proposed classifier in the {\tt sparsediscrim} R package,
available on CRAN.

\end{abstract}

<<knitr_setup, include=FALSE, echo=FALSE, cache=FALSE>>=
# Round digits throughout document including \Sexpr statements.
# See Yihui's response here:
# http://stackoverflow.com/questions/11062497/how-to-avoid-using-round-in-every-sexpr
options(digits=3)

opts_chunk$set(fig.align='default', dev='png', message=FALSE, warning=FALSE,
               cache=TRUE, echo=FALSE, autodep=TRUE, fig.width=12,
               fig.height=12, out.width="\\linewidth")
@

<<setup>>=
setwd("..")
library(ProjectTemplate)
load.project()
@

<<munge_contaminated_sims>>=
# Contaminated Normal Sims
results_contaminated <- lapply(results_block_contaminated[[1]], function(r) {
  with(r, c(error_rates, p=p, mu=mu, contamination_prob=contamination_prob))
})
results_contaminated <- do.call(rbind.data.frame, results_contaminated)
results_contaminated <- tbl_df(results_contaminated)

results_contaminated <- results_contaminated %>%
  group_by(p, contamination_prob, mu) %>%
  summarise_each(funs(mean)) %>%
  arrange(p, contamination_prob, mu) %>%
  ungroup

results_contaminated <- results_contaminated %>%
  gather(Classifier, Error, -p, -contamination_prob, -mu) %>%
  mutate(p=factor(p)) %>%
  filter(mu == 0.5)

# Remove underscore from classifier names
levels(results_contaminated$Classifier) <- gsub("_", " ",
                                                levels(results_contaminated$Classifier))

# Maximum Standard Error for Contaminated Sims
max_sim_se <- with(results_contaminated, max(sqrt(Error * (1 - Error) / 500)))
max_sim_se <- round(max_sim_se, digits=3)
@

<<munge_microarray_results>>=
# Microarray Classification Results
results_microarray <- lapply(results[[1]], function(r) {
  with(r, c(error_rates, `data_set`=data_set))
})
results_microarray <- do.call(rbind.data.frame, results_microarray)
results_microarray <- tbl_df(results_microarray)

avg_error_microarray <- results_microarray %>%
    group_by(data_set) %>%
    summarise_each(funs(mean)) %>%
    gather(Classifier, Avg_Error, -data_set)

sd_error_microarray <- results_microarray %>%
  group_by(data_set) %>%
  summarise_each(funs(sd)) %>%
  gather(Classifier, SD_Error, -data_set)

# Microarray classification results are presented in a table
# Rows = classifier, columns = data set
# Cells = Error (SD) rounded to 3 digits
results_microarray <- inner_join(avg_error_microarray, sd_error_microarray,
                                 by=c('data_set', 'Classifier')) %>%
  mutate(Table_Cells=sprintf('%.3f (%.3f)', Avg_Error, SD_Error),
         data_set=stringr::str_to_title(data_set),
         Classifier=gsub("_", " ", Classifier)) %>%
  select(-Avg_Error, -SD_Error) %>%
  spread(data_set, Table_Cells)
@

\section{Introduction}

As machine-learning algorithms are increasingly utilized to automate data-driven
decisions, researchers continue to develop a wealth of \emph{classifiers}
employed in a variety of disciplines, such as medical research, facial
recognition, remote sensing, and chemical analysis
\citep*{Ching:2012fu,Li:2012ev,Zhang:2010va}. Despite the advances in
classification to facilitate technological advancements, well-established
classifiers often degrade in accuracy with modern, high-dimensional data sets
when the number of features $p$ exceeds the total sample size $N$, due to the
\emph{curse of dimensionality} \citep{Bellman:1961tn}. For instance, linear
discriminant analysis (\emph{LDA}) and quadratic discriminant analysis
(\emph{QDA}) are popular because of their simplicity, speed of computation, and
interpretability. However, both methods are incalculable when $p > N$ because
the class sample covariance matrices are singular \citep*{Bouveyron:2007gx}.

As a result, researchers have emphasized three standard approaches to modifying
the \emph{LDA} and \emph{QDA} classifiers for application to high-dimensional
data: 1) covariance-matrix regularization, 2) dimension reduction, and 3)
parameter restrictions. Covariance-matrix regularization methods are often based
on the well-known ridge-regression approach of \cite{Hoerl:1970cd}.
\cite{Ching:2012fu}, \cite{Zhang:2010va}, \cite{Ye:2009gd}, \cite{Ji:2008wp},
\cite*{Guo:2007te}, \cite*{Srivastava:2007ud}, \cite{Ye:2006vx}, and
\cite{Ye:2005uu} have proposed regularization methods to improve the estimation
of the class covariance matrices by adjusting the eigenvalues of the sample
covariance matrices to ensure positive definiteness. Alternatively, dimension
reduction typically transforms the original data in the high-dimensional space
to a subspace having a much smaller dimension such that redundant and noisy
information is removed
\citep*{Fan:2012iq,Ye:2009gd,Bouveyron:2007gx,Ye:2006vx,Ye:2005hd,Antoniadis:2003bk}.
Finally, \cite*{Mai:2012bf}, \cite{Fan:2012iq}, \cite*{Merchante:2012vk},
\cite*{Clemmensen:2011kr}, \cite*{Huang:2010ju}, and \cite*{Dudoit:2002ev} have
proposed restrictions on the population covariance matrices to reduce the number
of parameters that must be estimated. For instance, \cite{Dudoit:2002ev} have
assumed the features within each class are independent so that the population
covariance matrices are diagonal, resulting in positive-definite sample
covariance matrices. For excellent overviews of the regularization problem in
discriminant analysis and other techniques, see \cite{Murphy:2012uq} and
\cite*{Mkhadri:1997gy}.

Here, we present a new classification method intended for the $p > N$ case that
builds on the classic regularized discriminant analysis (\emph{RDA}) proposed by
\cite{Friedman:1989tm}. Our proposed high-dimensional \emph{RDA} (\emph{HDRDA})
classifier employs a biased covariance matrix estimator that partially pools the
individual sample covariance matrices from the \emph{QDA} classifier with the
pooled sample covariance matrix from the \emph{LDA} classifier. We then shrink
the resulting covariance-matrix estimator towards a scaled identity matrix to
ensure positive definiteness. By reparameterizing the \emph{RDA} classifier
similar to \cite*{Hastie:2008dt} and \cite{Halbe:2007ea}, we first show that the
pooling parameter in the \emph{HDRDA} classifier determines the contribution of
each training observation to the estimation of each class covariance matrix,
enabling interpretability that has been previously lacking with the \emph{RDA}
classifier \citep{Bensmail:1996fw}.

Next, we achieve a substantial reduction in dimension to reduce greatly the
computation during model selection by applying a reasoning similar to
\cite{Ye:2006tq}. We establish that the matrix operations corresponding to the
null space of the pooled sample covariance matrix are redundant and can be
discarded from the \emph{HDRDA} decision rule without loss of classificatory
information. Moreover, because the \emph{HDRDA} decision rule is invariant to
adjustments to the zero eigenvalues, the decision rule in the original feature
space is equivalent to a decision rule in a lower dimension, such that matrix
inverses and determinants of relatively small matrices can be rapidly computed.
Furthermore, we show that several shrinkage methods that are special cases of
the \emph{HDRDA} classifier have no effect on the zero eigenvalues of the
covariance-matrix estimators. Such methods include work from
\cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and other methods studied by
\cite{Ramey:2013ji} and \citet*{Xu:2009fl}.

We provide an efficient algorithm that expedites the model-selection process for
the \emph{HDRDA} classifier when $p \gg N$. We then study the classification
performance of the \emph{HDRDA} classifier on four real high-dimensional data
sets along with a simulation design that generalizes the experiments initially
conducted by \cite{Guo:2007te}. We demonstrate that the \emph{HDRDA} classifier
often attains superior classification accuracy to several recent classifiers
designed for small-sample, high-dimensional data from \cite*{Tong:2012hw},
\cite{Witten:2011kc}, \cite*{Pang:2009ik}, and \cite{Guo:2007te}. We also
include as a benchmark the random forest from \cite{Breiman:2001fb} because
\cite*{FernandezDelgado:2014ul} has concluded that the random forest is often
superior to other classifiers in benchmark studies.

The remainder of this document is organized as follows. In Section
\ref{sec:preliminaries} we present the discrimination problem and the necessary
notation to describe our contributions. In Section \ref{sec:rda} we present the
\emph{HDRDA} classifier along with its interpretation, properties, and
computationally efficient model-selection procedure. Section \ref{sec:sims}
describes our simulation studies of artificial and real data sets and also
describes the experimental results. We conclude with a brief discussion in
Section \ref{sec:discussion}.

\section{Preliminaries}
\label{sec:preliminaries}

\subsection{Notation}

To facilitate our discussion of covariance-matrix regularization and regularized
discriminant analysis, we require the following notation. Let $\mathbb{R}_{a
  \times b}$ denote the matrix space of all $a \times b$ matrices over the real
field $\mathbb{R}$. Denote by $\bm I_m$ the $m \times m$ identity matrix, and
let $\bm 0_{m \times p}$ be the $m \times p$ matrix of zeros, such that $\bm
0_m$ is understood to denote $\bm 0_{m \times m}$. Define $\bm 1_m \in
\mathbb{R}_{m \times 1}$ as a vector of ones. Let $\bm A \oplus \bm B$ denote
the direct sum of $\bm A \in \mathbb{R}_{r \times r}$ and $\bm B \in
\mathbb{R}_{s \times s}$ \citep[Chapter 1]{Lutkepohl:1996uz}. Let $\bm A^{T}$,
$\bm A^+$, and $\mathcal{N}(\bm A)$ denote the transpose, the Moore-Penrose
pseudoinverse, and the null space of $\bm A \in \mathbb{R}_{m \times p}$,
respectively. Denote by $\mathbb{R}_{p \times p}^{>}$ the cone of real $p \times
p$ positive-definite matrices. Similarly, let $\mathbb{R}_{p \times p}^{\ge}$
denote the cone of real $p \times p$ positive-semidefinite matrices. Let
$V^{\perp}$ denote the orthogonal complement of a vector space $V \subset
\mathbb{R}_{p \times 1}$. For $c \in \mathbb{R}$, let $c^+ = 1/c$ if $c \ne 0$
and $0$ otherwise.

\subsection{Discriminant Analysis}

In discriminant analysis we wish to assign an unlabeled vector $\bm x \in
\mathbb{R}_{p \times 1}$ to one of $K$ unique, known classes by constructing a
classifier from $N$ training observations. Let $\bm x_i = (x_{i1}, \ldots,
x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation $(i = 1, \ldots,
N)$ with true, unique membership $y_i \in \{\omega_1, \ldots, \omega_K\}$. We
assume that $(\bm x_i, y_i)$ is a realization from a mixture distribution $p(\bm
x) = \sum_{k=1}^K p(\bm x | \omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$
is the probability density function (PDF) of the $k$th class and $p(\omega_k)$
is the prior probability of class membership of the $k$th class. We further
assume $p(\omega_k) = p(\omega_l)$, $1 \le k, l \le K$, $k \ne l$.

The \emph{QDA} classifier is the optimal Bayesian decision rule with respect to
a $0-1$ loss function when $p(\bm x | \omega_k)$ is the PDF of the multivariate
normal distribution with known mean vectors $\bm \mu_k \in \mathbb{R}_{p \times 1}$
and known covariance matrices $\bm \Sigma_k \in \mathbb{R}_{p \times p}^{>}$,
$k = 1, 2, \ldots, K$. Because $\bm \mu_k$ and $\bm \Sigma_k$ are typically
unknown, we assign an unlabeled observation $\bm x$ to class $\omega_k$ with the
sample \emph{QDA} classifier
\begin{align}
	D_{QDA}(\bm x) = \argmin_{k} (\bm x - \xbar_k)^{T}\widehat{\bm
    \Sigma}_k^{-1}(\bm x - \xbar_k) + \log |\widehat{\bm \Sigma}_k|, \label{eq:qda}
\end{align}
where $\xbar_k$ and $\widehat{\bm \Sigma}_k$ are the maximum likelihood
estimators (MLEs) of $\bm \mu_k$ and $\bm \Sigma_k$, respectively. If we assume
further that $\bm \Sigma_k = \bm \Sigma$, $k = 1, \ldots, K$, then the pooled
sample covariance matrix $\widehat{\bm \Sigma}$ is substituted for $\widehat{
  \bm \Sigma}_k$ in \eqref{eq:qda}, where
\begin{align}
	\widehat{\bm \Sigma} = N^{-1} \sum_{k=1}^K n_k \widehat{\bm \Sigma}_k \label{eq:pooled-cov}
\end{align}
is the MLE for $\bm \Sigma$. Here, \eqref{eq:qda} reduces to the sample
\emph{LDA} classifier. We omit the log-determinant because it is constant across
the $K$ classes.

The smallest eigenvalues of $\widehat{\bm \Sigma}_k$ and the directions
associated with their eigenvectors can highly influence \eqref{eq:qda}. In fact,
the eigenvalues of $\widehat{ \bm \Sigma}_k$ are well-known to be biased if $p
\ge n_k$ such that the smallest eigenvalues are underestimated
\citep{Seber:2004uh}. Moreover, if $p > n_k$, then rank$(\widehat{\bm \Sigma}_k)
\le n_k$, which implies that at least $p - n_k$ eigenvalues of $\widehat{\bm
  \Sigma}_k$ are zero. Furthermore, although more feature information is
available to discriminate among the $K$ classes, if $p > n_k$, \eqref{eq:qda} is
incalculable because $\widehat{\bm \Sigma}_k^{-1}$ does not exist.

Several regularization methods, such as the methods considered by
\cite{Xu:2009fl}, \cite{Guo:2007te}, and \cite{Mkhadri:1995jp} have been
proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$
so that \eqref{eq:qda} is calculable and has reduced variability in the standard
bias-variance tradeoff. A common form of the covariance-matrix regularization
applies a shrinkage factor $\gamma > 0$, so that
\begin{align}
	\widehat{\bm \Sigma}_k(\gamma) = \widehat{\bm \Sigma}_k + \gamma \bm I_p, \label{eq:ridge-estimator}
\end{align}
similar to a method employed in ridge regression \citep{Hoerl:1970cd}. Equation
\eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance
matrix $\widehat{\bm \Sigma}_k$ toward $\bm I_p$, thereby increasing the
eigenvalues of $\widehat{\bm \Sigma}_k$ by $\gamma$. Specifically, the zero
eigenvalues are replaced with $\gamma$, so that \eqref{eq:ridge-estimator} is
positive definite. For additional covariance-matrix regularization methods, see
\cite{Ramey:2013ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.

\section{High-Dimensional Regularized Discriminant Analysis}
\label{sec:rda}

Here, we define the \emph{HDRDA} classifier by first formulating the
covariance-matrix estimator $\widehat{\bm \Sigma}_k(\lambda)$ and demonstrating
its clear interpretation as a linear combination of the crossproducts of the
training observations centered by their respective class sample means. We define
the convex combination
\begin{align}
  \widehat{\bm \Sigma}_k(\lambda) \coloneqq (1 - \lambda) \widehat{\bm \Sigma}_k + \lambda
  \widehat{\bm \Sigma}, \quad k = 1, \ldots, K,\label{eq:sig-lambda-alternative}
\end{align}
where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. By rewriting
\eqref{eq:sig-lambda-alternative} in terms of the observations $ \bm x_i$ $(i =
1, \ldots, N)$, each centered by its class sample mean, we attain a clear
interpretation of $\widehat{\bm \Sigma}_k(\lambda)$. That is,
\begin{align}
	\widehat{\bm \Sigma}_k(\lambda)
  &= \left( 1 - \lambda + \frac{\lambda n_k}{N} \right) \widehat{\bm \Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}}^K n_{k'} \widehat{\bm \Sigma}_{k'} \nonumber \\
	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N
  I(y_i = k) \bm x_i \bm x_i^{T} +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k)
  \bm x_i \bm x_i^{T} \nonumber \\
	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i^{T},\label{eq:sig-lambda-alternative2}
\end{align}
where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$.
From \eqref{eq:sig-lambda-alternative2}, $\lambda$ weights the contribution of
each of the $N$ observations in estimating $\bm \Sigma_k$ from all $K$ classes
rather than $n_k$ observations from a single class. As a result, we can
interpret \eqref{eq:sig-lambda-alternative2} as a covariance-matrix estimator
that borrows from $\widehat{\bm \Sigma}$ in \eqref{eq:pooled-cov} to estimate $
\bm \Sigma_k$.

Below, we will show that the pooling operation is advantageous in increasing the
rank of each $\widehat{\bm \Sigma}_k(\lambda)$ from rank$(\widehat{\bm
  \Sigma}_k)$ to rank$(\widehat{\bm \Sigma})$ for $0 < \lambda \le 1$. Notice
that if $\lambda = 0$, then the observations from the remaining $K - 1$ classes
do not contribute to the estimation of $\bm \Sigma_k$, corresponding to
$\widehat{\bm \Sigma}_k$. Furthermore, if $\lambda = 1$, the weights
$c_{ik}(\lambda)$ in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$,
corresponding to $\widehat{\bm \Sigma}$. For brevity, when $\lambda = 1$, we
define $\bm X = [\sqrt{c_{1k}(1)} \bm x_1^{T}, \ldots, \sqrt{c_{Nk}(1)} \bm
  x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma} = N^{-1} \bm X^{T} \bm X$. Similarly,
for $\lambda = 0$, we define $\bm X_k = [\sqrt{c_{1k}(0)} \bm x_1^{T}, \ldots,
  \sqrt{c_{Nk}(0)} \bm x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma}_k = n_k^{-1} \bm
X_k^{T} \bm X_k$. In Figure \ref{fig:hdrda-contours} we plot the contours of
five multivariate normal populations for $\lambda = 0$ with unequal covariance
matrices. As $\lambda$ approaches 1, the contours become more similar, resulting
in identical contours for $\lambda = 1$.

\begin{figure}
<<hdrda_contours, fig.width=8, fig.height=8, out.width='.33\\linewidth', dpi=200, fig.show='hold'>>=
rda_contours(lambda = 0)
rda_contours(lambda = 0.2)
rda_contours(lambda = 0.4)
rda_contours(lambda = 0.6)
rda_contours(lambda = 0.8)
rda_contours(lambda = 1)
@
\caption{Contours of five multivariate normal populations as a function of the
  pooling parameter $\lambda$.}
\label{fig:hdrda-contours}
\end{figure}

As we have discussed above, several eigenvalue adjustment methods have been
proposed to increase eigenvalues (approximately) equal to 0. To further improve
the estimation of $\bm \Sigma_k$ and to stabilize the estimator's inverse, we
define the eigenvalue adjustment of \eqref{eq:sig-lambda-alternative} as
\begin{align}
	\tilde{\bm \Sigma}_k \coloneqq \alpha_k \widehat{\bm \Sigma}_k(\lambda) + \gamma \bm I_p,\label{eq:hdrda-cov}
\end{align}
where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage
constant. Thus, the \emph{pooling} parameter $\lambda$ controls the amount
borrowed from $\widehat{\bm \Sigma}$ to estimate $\bm \Sigma_k$, and the
\emph{shrinkage} parameter $\gamma$ determines the degree of eigenvalue
shrinkage. The choice of $\alpha_k$ allows for a flexible formulation of
covariance-matrix estimators. For instance, if $\alpha_k = 1$, $k = 1, \ldots,
K$, then \eqref{eq:hdrda-cov} resembles \eqref{eq:ridge-estimator}. Similarly,
if $\alpha_k = 1 - \gamma$, then \eqref{eq:hdrda-cov} has a form similar to the
\emph{RDA} classifier from \cite{Friedman:1989tm}. Substituting
\eqref{eq:hdrda-cov} into \eqref{eq:qda}, we define the \emph{HDRDA} classifier
\begin{align}
	D_{HDRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)^{T}\tilde{\bm \Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm \Sigma}_k|. \label{eq:hdrda}
\end{align}
For $\gamma > 0$, $\tilde{\bm \Sigma}_k$ is nonsingular such that $\tilde{\bm
  \Sigma}_k^{-1}$ can be substituted for $\tilde{\bm \Sigma}_k^{+}$ in
\eqref{eq:hdrda}. If $\gamma = 0$, we explicitly set $|\tilde{\bm \Sigma}_k|$
equal to the product of the positive eigenvalues of $\tilde{\bm
  \Sigma}_k$. Following \cite{Friedman:1989tm}, we select $\lambda$ and $\gamma$
from a grid of candidate models via cross-validation. We provide an
implementation of \eqref{eq:hdrda} in the {\tt hdrda} function in the {\tt
  sparsediscrim} R package, available on CRAN.

The choice of $\alpha_k$ in \eqref{eq:hdrda-cov} is one of convenience and
allows the flexibility of various covariance-matrix estimators proposed in the
literature. In practice, we generally are not interested in estimating
$\alpha_k$ as it would incorporate $K$ additional tuning parameters to estimate
via cross-validation, which is counterproductive to our goal of improving the
\emph{RDA} classifier's efficiency. By setting $\alpha_k$ to an appropriate
constant, the \emph{HDRDA} covariance-matrix estimator includes or resembles a
large family of estimators. Notice that if $\alpha_k = 1$ and $\lambda = 1$,
\eqref{eq:hdrda-cov} is equivalent to the standard ridge-like covariance-matrix
estimator in \eqref{eq:ridge-estimator}, and various estimators proposed in the
literature can be obtained by selecting $\gamma$ accordingly. For instance, with
$\gamma = \tr\{ \widehat{\bm \Sigma} \} / \min(N, p)$, we obtain the estimator
from \cite{Srivastava:2007ww}.

Perhaps the most well-known of the ridge-like estimators is from
\cite{Friedman:1989tm}, who proposed regularized discriminant analysis (RDA) by
incorporating a weighted average of the sample covariance matrix $\widehat{ \bm
  \Sigma}_k$ and the pooled sample covariance matrix $\widehat{\bm \Sigma}$ to
estimate $\bm \Sigma_k$ with
\begin{align}
    \widehat{\bm \Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda) \quad
    (k =
      1, \ldots, K),\label{eq:sig-lambda}
      \end{align}
where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda N$, $
\bm S_k = n_k \widehat{\bm \Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, and $ \bm
S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$. If $\lambda = 0$,
\eqref{eq:sig-lambda} is equal to $\widehat{\bm \Sigma}_k$ in \eqref{eq:qda}.
Alternatively, if $\lambda = 1$, then \eqref{eq:sig-lambda} reduces to
\eqref{eq:pooled-cov}. To further improve the estimation of $\bm \Sigma_k$ and
to stabilize the inverse of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm} has
proposed the biased covariance-matrix estimator
\begin{align}
          \widehat{\bm \Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm
            \Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\bm
              \Sigma}_k(\lambda)\}}{p} \bm I_p,\label{eq:sig-rda}
          \end{align}
where $\gamma \in [0, 1]$ is a regularization parameter that controls the
shrinkage of \eqref{eq:sig-rda} towards $\bm I_p$ and is weighted by the average
of the eigenvalues of \eqref{eq:sig-lambda}. Finally, the RDA classifier is
established by substituting \eqref{eq:sig-rda} into \eqref{eq:qda}.

Our proposed \emph{HDRDA} classifier resembles \eqref{eq:sig-rda} but improves
the interpretation of the pooling parameter $\lambda$ as discussed
above. Moreover, as we discuss below, our formulation of \emph{HDRDA} classifier
lends itself to rapid estimation in high-dimensional classification.

\section{Properties of the HDRDA Classifier}
\label{sec:hdrda-properties}

Next, we establish properties of the covariance-matrix estimator and the
decision rule employed in the \emph{HDRDA} classifier. By doing so, we
demonstrate that \eqref{eq:hdrda} lends itself to a more efficient
calculation. We decompose \eqref{eq:hdrda} into a sum of two components, where
the first summand consists of matrix operations applied to low-dimensional
matrices and the second summand corresponds to the null space of $\widehat{\bm
  \Sigma}$ in \eqref{eq:pooled-cov}. We show that the matrix operations
performed on the null space of $\widehat{\bm \Sigma}$ yield constant quadratic
forms across all classes and can be omitted. For $p \gg N$, the constant
component involves determinants and inverses of high-dimensional matrices, and
by ignoring these calculations, a substantial reduction in computational costs
is achieved. Furthermore, a byproduct is that adjustments to the associated
eigenvalues have no effect on \eqref{eq:hdrda}. Lastly, we utilize the singular
value decomposition to efficiently calculate the eigenvalue decomposition of
$\widehat{\bm \Sigma}$, further reducing the computational costs of the
\emph{HDRDA} classifier.

First, we require the following relationship regarding the null spaces of
$\widehat{\bm \Sigma}_k(\lambda)$, $\widehat{\bm \Sigma}$, and $\widehat{ \bm
  \Sigma}_k$.

\begin{lemma}\label{lemma:null-spaces}
Let $\widehat{\bm \Sigma}_k$ and $\widehat{\bm \Sigma}$ be the MLEs of $\bm
\Sigma_k$ and $\bm \Sigma$, respectively. Let $\widehat{\bm \Sigma}_k(\lambda)$
be defined as in \eqref{eq:sig-lambda-alternative}. Then,
$\mathcal{N}\{\widehat{\bm \Sigma}_k(\lambda)\} \subset \mathcal{N}(\widehat{
  \bm \Sigma}) \subset \mathcal{N}(\widehat{\bm \Sigma}_k)$, $k = 1, \ldots, K$.
\end{lemma}
\begin{proof}
Let $\bm z \in \mathcal{N}\{\widehat{\bm \Sigma}_k(\lambda)\}$ for some $k = 1,
\ldots, K$. Hence, $0 = \bm z^{T} \widehat{\bm \Sigma}_k(\lambda) \bm z = (1 -
\lambda) \bm z^{T} \widehat{\bm \Sigma}_k \bm z + \lambda \bm z^{T} \widehat{
  \bm \Sigma} \bm z$. Because $\widehat{\bm \Sigma}_k, \widehat{\bm \Sigma}\in
\mathbb{R}_{p \times p}^{\ge}$, we have $\bm z \in \mathcal{N}(\widehat{
  \bm \Sigma})$ and $\bm z \in \mathcal{N}(\widehat{\bm \Sigma}_k)$. In particular,
we have that $\mathcal{N}\{\widehat{\bm \Sigma}_k(\lambda)\} \subset
\mathcal{N}(\widehat{\bm \Sigma})$. Now, suppose $\bm z \in
\mathcal{N}(\widehat{\bm \Sigma})$. Similarly, we have that $0 = \bm z^{T}
\widehat{\bm \Sigma} \bm z = N^{-1} \sum_{k = 1}^K n_k \bm z^{T} \widehat{
  \bm \Sigma}_k \bm z$, which implies that $\bm z \in \mathcal{N}(\widehat{
  \bm \Sigma}_k)$ because $\widehat{\bm \Sigma}_k \in \mathbb{R}_{p \times
  p}^{\ge}$. Therefore, $\mathcal{N}(\widehat{\bm \Sigma}) \subset
\mathcal{N}(\widehat{\bm \Sigma}_k)$.
\end{proof}

In Lemma \ref{lemma:rda-tilde-Sigma_k} below, we derive an alternative
expression for $\tilde{\bm \Sigma}_k$ in terms of the matrix of eigenvectors of
$\widehat{\bm \Sigma}$. Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U^{T}$ be the
eigendecomposition of $\widehat{\bm \Sigma}$ such that $\bm D \in \mathbb{R}_{p
  \times p}^{\ge}$ is the diagonal matrix of eigenvalues of $\widehat{\bm \Sigma}$
with $\bm D = \bm D_q \oplus \bm 0_{p-q}$, $\bm D_q \in \mathbb{R}_{q \times
  q}^{>}$ is the diagonal matrix consisting of the positive eigenvalues of
$\widehat{\bm \Sigma}$, the columns of $\bm U \in \mathbb{R}_{p \times p}$ are
the corresponding orthonormal eigenvectors of $\widehat{\bm \Sigma}$, and
rank$(\widehat{\bm \Sigma}) = q$. Then, we partition $\bm U = (\bm U_1,
\bm U_2)$ such that $\bm U_1 \in \mathbb{R}_{p \times q}$ and $\bm U_2 \in
\mathbb{R}_{p \times (p - q)}$.

\begin{lemma}\label{lemma:rda-tilde-Sigma_k}
Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U^{T}$ be the eigendecomposition of
$\widehat{\bm \Sigma}$ as above, and suppose that rank$(\widehat{\bm \Sigma}) =
q \le p$. Then, we have
\begin{align}
	\tilde{\bm \Sigma}_k &= \bm U(\bm W_k \oplus \gamma \bm I_{p-q})\bm U^{T}, \quad
  k = 1, \ldots, K,\label{eq:rda-matrix}
\intertext{where}
\bm W_k &= \alpha_k \{(1 - \lambda) \bm U_1^{T} \widehat{\bm \Sigma}_k \bm U_1 + \lambda \bm D_q\} + \gamma \bm I_{q}.\label{eq:Wk}
\end{align}
\end{lemma}
\begin{proof}
From Lemma \ref{lemma:null-spaces}, the columns of $\bm U_2$ span the null space
of $\widehat{\bm \Sigma}_k$, which implies that $\widehat{\bm \Sigma}_k \bm U_2
= \bm 0_{p \times (p - q)}$. Hence, $\bm U^{T} \widehat{\bm \Sigma}_k \bm U =
\bm U_1^{T} \widehat{\bm \Sigma}_k \bm U_1 \oplus \bm 0_{p-q}$, $k = 1, \ldots,
K$. Thus, $\bm U^{T} \tilde{\bm \Sigma}_k \bm U = \alpha_k \{(1 - \lambda) \bm
U^{T} \widehat{\bm \Sigma}_k \bm U + \lambda \bm D\} + \gamma \bm I_p$, and
\eqref{eq:rda-matrix} holds because $\bm U$ is orthogonal.
\end{proof}

As an immediate immediate consequence of Lemma \ref{lemma:rda-tilde-Sigma_k}, we
have the following corollary.

\begin{cor}
Let $\widehat{\bm \Sigma}_k(\lambda)$ be defined as in
\eqref{eq:sig-lambda-alternative}. Then, for $\lambda \in (0, 1]$,
  rank$\{\widehat{\bm \Sigma}_k(\lambda)\} = q$, $k = 1, \ldots, K$.
\end{cor}
\begin{proof}
The proof follows by setting $\gamma = 0$ in Lemma \ref{lemma:rda-tilde-Sigma_k}.
\end{proof}

Thus, by incorporating each $\bm x_i$ into the estimation of $\bm \Sigma_k$, we
increase the rank of $\widehat{\bm \Sigma}_k(\lambda)$ to $q \approx N$ if
$\lambda \ne 0$. Next, we provide an essential result that enables us to prove
that \eqref{eq:hdrda} is invariant to adjustments to the eigenvalues of $\tilde{
  \bm \Sigma}_k$ corresponding to the null space of $\widehat{\bm \Sigma}$.

\begin{lemma}\label{lemma:RDA-constant-term}
Let $\bm U_2$ be defined as above. Then, for all $\bm x \in \mathbb{R}_{p \times
  1}$, $\bm U_2^{T} (\bm x - \xbar_k) = \bm U_2^{T} (\bm x - \xbar_{k'})$, $1
\le k, k' \le K$, where $k \ne k'$.
\end{lemma}
\begin{proof}
Let $\bm x \in \mathbb{R}_{p \times 1}$, and suppose that $1 \le k, k' \le K$.
Recall that $\bm U_2 \in \mathcal{N}(\widehat{\bm \Sigma})$, which implies that
$\bm U_2^{T} \in \mathcal{C}(\widehat{\bm \Sigma})^{\perp}$ \citep[Lemma
  1.2.5]{Kollo:2005vp}. Now, because $\bm x_i \in \mathcal{C}(\widehat{ \bm
  \Sigma})$ $(i = 1, \ldots, N)$, $\bm U_2^{T} \bm x_i = \bm 0_{p-q}$.  Hence,
$\bm 0_{p-q} = \sum_{i=1}^N \beta_i \bm U_2^{T}\bm x_i = \bm U_2^{T}(\xbar_k -
\xbar_{k'})$, where $\beta_i = (n_k n_{k'})^{-1} \{ I(y_i = k) n_{k'} - I(y_i =
k') n_k \}$. Therefore, $\bm U_2^{T} (\bm x - \xbar_k) = \bm U_2^{T} (\bm x -
\xbar_{k'})$.
\end{proof}

We now present our main result where we decompose \eqref{eq:hdrda} and show that
the term requiring the largest computational costs does not contribute to the
classification of an unlabeled observation performed using Lemma
\ref{lemma:RDA-constant-term}. Hence, we reduce \eqref{eq:hdrda} to an
equivalent, more computationally efficient decision rule.

\begin{thm}
Let $\tilde{\bm \Sigma}_k$ and $\bm W_k$ be defined as in \eqref{eq:rda-matrix}
and \eqref{eq:Wk}, respectively, and let $\bm U_1$ be defined as above. Then,
the decision rule in \eqref{eq:hdrda} is equivalent to
	\begin{align}
		D_{HDRDA}(\bm x) &= \argmin_k  (\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k) + \log | \bm W_k |. \label{eq:hdrda-decomposed}
	\end{align}
\end{thm}
\begin{proof}
From \eqref{eq:rda-matrix}, we have that $\tilde{\bm \Sigma}_k^{+} = \bm U(
\bm W_k^{-1} \oplus \gamma^{+} \bm I_{p-q} )\bm U^{T}$ and $|\tilde{\bm \Sigma}_k| =
\gamma^{p-q}| \bm W_k |$, $k = 1, \ldots, K$. Therefore, for all $\bm x \in
\mathbb{R}_{p \times 1}$, we have that
	\begin{align*}
	(\bm x - \xbar_k)^{T} \tilde{\bm \Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm \Sigma}_k| &= (\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k)\\
	&+ \gamma^{+} (\bm x - \xbar_k)^{T} \bm U_2 \bm U_2^{T} (\bm x - \xbar_k) + \log | \bm W_k |\\
  &+ (p - q) \log \gamma.
	\end{align*}
Because $\gamma$ is constant for $k = 1, \ldots, K$, we can omit the $(p - q)
\log \gamma$ term and particularly avoid the calculation of $\log 0$ for $\gamma
= 0$. Then, the proof follows from Lemma \ref{lemma:RDA-constant-term} because
$\bm U_2^{T} (\bm x - \xbar_k)$ is constant for $k = 1, \ldots, K$.
\end{proof}

As we can see, we are able to avoid the time-consuming inverses and determinants
of $p \times p$ covariance matrices in \eqref{eq:hdrda} and instead calculating
these same operations on $\bm W_k \in \mathbb{R}_{q \times q}$ in
\eqref{eq:hdrda-decomposed}. The substantial computational improvements arise
because our proposed classifier in \eqref{eq:hdrda} is invariant to the term
$\bm U_2$, thus yielding an equivalent classifier in \eqref{eq:hdrda-decomposed}
with a substantial reduction in computational complexity. Here, we demonstrate
that the computational efficiency in calculating the inverse and determinant of
$\bm W_k$ can be further improved via standard matrix operations by showing that
the inverses and determinants of $\bm W_k$ can be performed on matrices of size
$n_k \times n_k$.

\begin{proposition}\label{proposition:hdrda-W_k}
Let $\bm W_k$ be defined as above. Then, $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$
and
	\begin{align}
		\bm W_k^{-1} &= \bm \Gamma_k^{-1} - n_k^{-1} \alpha_k(1 - \lambda) \bm \Gamma_k^{-1} \bm U_1^{T} \bm X_k^{T} \bm Q_k^{-1} \bm X_k \bm U_1 \bm \Gamma_k^{-1},\label{eq:W_k_inv}
    \intertext{where}
    \bm Q_k &= \bm I_{n_k} + n_k^{-1} \alpha_k(1 - \lambda) \bm X_k \bm U_1 \bm \Gamma_k^{-1} \bm U_1^{T} \bm X_k^{T}\label{eq:Q_k}
    \intertext{and}
    \bm \Gamma_k &= \alpha_k \lambda \bm D_q + \gamma \bm I_q\label{eq:Gamma_k}.
	\end{align}
\end{proposition}
\begin{proof}
First, we write $\bm W_k = n_k^{-1} \alpha_k (1 - \lambda) \bm U_1^{T} \bm
X_k^{T} \bm X_k \bm U_1 + \bm \Gamma_k$. To calculate $|\bm W_k|$, we apply
Theorem 18.1.1 from \cite{Harville:2008wja}, which states that $|\bm A + \bm B
\bm T \bm C| = |\bm A| |\bm T| |\bm T^{-1} + \bm C \bm A^{-1} \bm B|$, where
$\bm A \in \mathbb{R}_{a \times a}^{>}$, $\bm B \in \mathbb{R}_{a \times b}$,
$\bm T \in \mathbb{R}_{b \times b}^{>}$, and $\bm C \in \mathbb{R}_{b \times
  a}$. Thus, setting $\bm A = \bm \Gamma_k$, $\bm B = \alpha_k (1 - \lambda) \bm
U_1^{T} \bm X_k^{T}$, $\bm T = \bm I_{n_k}$, and $\bm C = \bm X_k \bm U_1$, we
have $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$. Similarly, \eqref{eq:W_k_inv}
follows from the well-known Sherman-Woodbury formula \citep[Theorem
  18.2.8]{Harville:2008wja} by noting that $(\bm A + \bm B \bm T \bm C)^{-1} =
\bm A^{-1} - \bm A^{-1} \bm B (\bm T^{-1} + \bm C \bm A^{-1} \bm B)^{-1} \bm C
\bm A^{-1}$.
\end{proof}

The application of the Sherman-Woodbury formula requires that $\bm \Gamma_k$ be
nonsingular. This assumption is not met when $(\lambda, \gamma) = (0, 0)$
because $\bm \Gamma_k = \bm 0_q$. Here, we use the formulation in
\eqref{eq:hdrda-decomposed} instead. Also, notice that if $\alpha_k$ is constant
across the $K$ classes, then $\bm \Gamma_k$ in \eqref{eq:Gamma_k} is independent
of $k$. Consequently, $|\bm \Gamma_k|$ is constant across the $K$ classes and
need not be calculated in \eqref{eq:hdrda-decomposed}.

\subsection{Model Selection}

Thus far, we have presented the \emph{HDRDA} classifier and its properties that
facilitate an efficient calculation of the decision rule. Here, we describe an
efficient model-selection procedure along with pseudocode in Algorithm
\ref{alg:hdrda_model_selection} to select the optimal tuning-parameter estimates
from the Cartesian product of candidate values $\{\lambda_g\}_{g=1}^G \times
\{\gamma_h\}_{h=1}^H$. We estimate the $V$-fold cross-validation error rate for
each candidate pair and select $(\widehat{\lambda}, \widehat{\gamma})$ that
attains the minimum error rate. To calculate the $V$-fold cross-validation, we
partition the original training data into $V$ mutually exclusive and exhaustive
folds that have approximately the same number of observations. Then, for $v = 1,
\ldots, V$, we classify the observations in the $v$th fold by training a
classifier on the remaining $V - 1$ folds. We then calculate the
cross-validation error as the proportion of misclassified observations across
the $V$ folds.

\IncMargin{1em}
\begin{algorithm}
  \SetKwInOut{Input}{input}
  \SetKwInOut{Output}{output}

  \Input{
    Data matrix $\bm X$\\
    Parameter grid $\{\lambda_g\}_{g=1}^G \times \{\gamma_h\}_{h=1}^H$
  }
  \Output{Optimal Estimates $(\hat{\lambda}, \hat{\gamma})$}
  \BlankLine
  \For{$v \leftarrow 1$ \KwTo $V$} {
    Partition $\bm X$ into $\bm X_{train} \in \mathbb{R}_{N \times p}$ and
    $\bm X_{test} \in \mathbb{R}_{N_T \times p}$

    \For{$k\leftarrow 1$ \KwTo $K$}{
      Extract $\bm X_k \in \mathbb{R}_{n_k \times p}$ from $\bm X_{train}$

      Compute sample mean $\xbar_k$ from $\bm X_k$

      Center $\bm X_k \leftarrow \bm X_k - \bm 1_{n_k} \xbar_k^T$
    }

    $\bm X_c \leftarrow [\bm X_1^T, \ldots, \bm X_K^T]^T$

    Compute the compact SVD $\bm X_c = \bm M_q \bm D_q \bm U_1^T$

    Transform $\bm X_c \leftarrow \bm X_c \bm U_1$

    Transform $\bm X_{test} \leftarrow \bm X_{test} \bm U_1$

    \For{$k \leftarrow 1$ \KwTo $K$} {
      Extract $\bm X_k \in \mathbb{R}_{n_k \times q}$ from $\bm X_c$

      Recompute sample mean $\xbar_k$ from $\bm X_k$
    }

    \For{$(\lambda, \gamma) \in \{\lambda_g\}_{g=1}^G \times \{\gamma_h\}_{h=1}^H$} {
      \For{$k \leftarrow 1$ \KwTo $K$} {
        Compute $\bm Q_k$ using \eqref{eq:Q_k}

        Compute $\bm \Gamma_k$ using \eqref{eq:Gamma_k}

        Compute $\bm W_k^{-1}$ using \eqref{eq:W_k_inv}

        Compute $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$

        Compute $(\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x -
        \xbar_k) + \log | \bm W_k |$ for each row $\bm x$ of $\bm X_{test}$
      }
      Classify test observations $\bm X_{test}$ using \eqref{eq:hdrda-decomposed}

      Compute the number of misclassified test observations $\#\{\text{Error}_v(\lambda, \gamma)\}$
    }
   }
  Compute $\widehat{\text{Error}}(\lambda, \gamma) = N^{-1} \sum_{v=1}^V \#\{\text{Error}_v(\lambda, \gamma)\}$

  Report optimal $(\hat{\lambda}, \hat{\gamma}) \leftarrow \argmin_{(\lambda, \gamma)} \widehat{\text{Error}}(\lambda, \gamma)$
  \caption{Model selection for the \emph{HDRDA} classifier}
  \label{alg:hdrda_model_selection}
\end{algorithm}
\DecMargin{1em}

A primary contributing factor to the efficiency of Algorithm
\ref{alg:hdrda_model_selection} is our usage of the compact singular value
decomposition (SVD). Rather than computing the eigenvalue decomposition of
$\widehat{\bm \Sigma}$ to obtain $\bm U_1$, we instead obtain $\bm U_1$ by
computing the eigendecomposition of a much smaller $N \times N$ matrix when $p
\gg N$ \cite[Chapter~18.3.5]{Hastie:2008dt}. Applying the SVD, we decompose $\bm
X_c = \bm M \bm \Delta \bm U^{T}$, where $\bm M \in \mathbb{R}_{N \times p}$ is
orthogonal, $\bm \Delta \in \mathbb{R}_{p \times p}^{\ge}$ is a diagonal matrix
consisting of the singular values of $\bm X_c$, and $\bm U \in \mathbb{R}_{p
  \times p}$ is orthogonal. Recalling that $\widehat{\bm \Sigma} = N^{-1} \bm
X_c^{T} \bm X_c$, we have the eigendecomposition $\widehat{\bm \Sigma} = \bm U
\bm D \bm U^{T}$, where $\bm U$ is the matrix of eigenvectors of $\widehat{\bm
  \Sigma}$ and $\bm D = N^{-1} \bm \Delta$ is the diagonal matrix of eigenvalues
of $\widehat{\bm \Sigma}$. Now, we can obtain $\bm M$ and $\bm D$ efficiently
from the eigenvalue decomposition of the $N \times N$ matrix $\bm X_c \bm
X_c^{T} = \bm M \bm D \bm M^{T}$. Next, we compute $\bm U = \bm X_c^{T} \bm M
\bm D^{+/2}$, where $\bm D^{+/2} = \bm D_q^{-1/2} \oplus \bm 0_{N-q}$. We then
determine $q$, the number of numerically nonzero eigenvalues present in $\bm D$,
by calculating the number of eigenvalues that exceeds some tolerance value, say,
$1 \times 10^{-6}$. We then extract $\bm U_1$ as the first $q$ columns of $\bm
U$.

As a result of the compact SVD, we need calculate $\bm X_c \bm U_1$ only once
per cross-validation fold, requiring $O(p q N) \approx O(p N^2)$ calculations.
Hence, the computational costs of expensive calculations, such as matrix
inverses and determinants, are greatly reduced because they are performed in the
$q$-dimensional subspace. Similarly, we reduce the dimension of the test data
set by calculating $\bm X_{test} \bm U_1$ once per fold. Conveniently, we see
that the most costly computation involved in $\bm Q_k$ and $\bm W_k^{-1}$ is
$\bm X_k \bm U_1$, which can be extracted from $\bm X_c \bm U_1$. Thus, after
the initial calculation of $\bm X_c \bm U_1$ per cross-validation fold, $\bm
Q_k$ requires $O(n_k q^2)$ operations. Because $\bm Q_k \in \mathbb{R}_{n_k
  \times n_k}$, both its determinant and inverse require $O(n_k^3)$
operations. Consequently, $\bm W_k^{-1}$ requires $O(n_k q^2)$ operations. Also,
the inverse of the diagonal matrix $\bm \Gamma_k^{-1} \in \mathbb{R}_{q \times
  q}$ requires $O(q)$ operations. Finally, $| \bm W_k |$ requires $O(n_k^3)$
operations.

The expressions given in Proposition \ref{proposition:hdrda-W_k} also expedite
the selection of $\lambda$ and $\gamma$ via cross-validation because the most
time-consuming matrix operation involved in computing $\bm W_k^{-1}$ and $|\bm
W_k|$ is $\bm X_k \bm U_1 \in \mathbb{R}_{n_k \times q}$, which is independent
of $\lambda$ and $\gamma$. The subsequent operations in calculating $\bm
W_k^{-1}$ and $|\bm W_k|$ can be simply updated for different pairs of $\lambda$
and $\gamma$ without repeating the costly computations. When calculating the
discriminant scores of each test observation in $\bm X_{test}$, rather than
calculating each score individually, we can more efficiently calculate the
quadratic forms via a single matrix operation. That is, rather than calculating
$(\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k)$
individually for each row $\bm x$ of $\bm X_{test}$, we can instead calculate
$(\bm X_{test} - \xbar_k \bm 1_k')' \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm
X_{test} - \xbar_k \bm 1_k')$. The diagonal elements of the resulting matrix
contain the individual quadratic form of each test observation, $\bm x_t$.

\section{Classification Study}
\label{sec:sims}

In this section, we compare our proposed classifier with four classifiers
recently proposed for small-sample, high-dimensional data along with the
random-forest classifier from \cite{Breiman:2001fb} using version 3.2.2 of the
open-source statistical software {\tt R}. Within our study, we included
penalized linear discriminant analysis from \cite{Witten:2011kc}, implemented in
the {\tt penalizedLDA} package. We also considered shrunken centroids
regularized discriminant analysis from \cite{Guo:2007te} in the {\tt rda}
package. Because the {\tt rda} package does not perform the authors' ``Min-Min''
rule automatically, we applied this rule within our {\tt R} code. We included
two modifications of diagonal linear discriminant analysis from
\cite{Tong:2012hw} and \cite{Pang:2009ik}, where the former employs an improved
mean estimator and the latter utilizes an improved variance estimator. Both
classifiers are available in the {\tt sparsediscrim} package. Finally, we
incorporated the random forest as a benchmark based on the findings of
\cite{FernandezDelgado:2014ul}, who concluded that the random forest is often
superior to other classifiers in benchmark studies. We used the implementation
of the random-forest classifier from the {\tt randomForest} package with 250
trees and 100 maximum nodes. For each classifer we explicitly set prior
probabilities as equal, if applicable. All other classifier options were set to
their default settings. Below, we refer to each classifier by the first author's
surname. All simulations were conducted on an Amazon Elastic Compute Cloud (EC2)
{\tt c4.4xlarge} instance.

For the \emph{HDRDA} classifier in \eqref{eq:hdrda-decomposed}, we examined the
classification performance of two models. For the first \emph{HDRDA} model, we
set $\alpha_k = 1$, $k = 1, \ldots, K$, so that the covariance-matrix estimator
\eqref{eq:hdrda-cov} resembled \eqref{eq:ridge-estimator}. We estimated
$\lambda$ from a grid of 21 equidistant candidate values between 0 and 1,
inclusively. Similarly, we estimated $\gamma$ from a grid consisting of the
values $10^{-1}, \ldots, 10^4$, and $10^5$. We selected optimal estimates of
$\lambda$ and $\gamma$ using $V$-fold cross-validation with $V = 10$
\citep{Hastie:2008dt}. For the second model, we set $\alpha_k = 1 - \gamma$, $k
= 1, \ldots, K$, to resemble \citeauthor{Friedman:1989tm}'s parameterization. We
estimated both $\lambda$ and $\gamma$ from a grid of 21 equidistant candidate
values between 0 and 1, inclusively.

\subsection{Simulation Study}

In this section we compare the competing classifiers using the simulation design
from \cite{Guo:2007te}. This design is widely used within the high-dimensional
classification literature including the studies by \cite{Ramey:2013ji} and
\cite{Witten:2011kc}. First, we consider the block-diagonal covariance matrix
from from \cite{Guo:2007te}
\begin{align}
  \bm\Sigma_k = \bm\Sigma^{(\rho_k)} \oplus \bm\Sigma^{(-\rho_k)} \oplus \ldots \oplus \bm\Sigma^{(\rho_k)},\label{eq:block-sigma}
\end{align}
where the $(i,j)$th entry of the block matrix $\bm\Sigma^{(\rho_k)} \in
\mathbb{R}_{100 \times 100}$ is
\begin{align*}
\bm\Sigma_{ij}^{(\rho_k)} = \{ \rho_k^{|i - j|} \}_{1 \le i,j \le 100}.
\end{align*}
The block-diagonal covariance structure in \eqref{eq:block-sigma} resembles
gene-expression data: within each block of pathways, genes are correlated, and
the correlation decays as a function of the distance between any two genes. The
original design from \cite{Guo:2007te} comprised two $p$-dimensional
multivariate normal populations with a common block-diagonal covariance matrix
where $\bm\Sigma_1 = \bm\Sigma_2$.

Although the design is indeed standard, the simulation configuration lacks
artifacts observed commonly in real data, such as skewness and extreme
outliers. As a result, we wished to investigate the effect of outliers on the
high-dimensional classifiers. To accomplish this goal, we generalized the
block-diagonal simulation configuration by sampling from a $p$-dimensional
multivariate contaminated normal distribution.  Denoting the PDF of the
$p$-dimensional multivariate normal distribution by $N_p(\bm x | \bm \mu, \bm
\Sigma)$, we write the PDF of the $k$th class as
\begin{align}
  p(\bm x | \omega_k) = (1 - \epsilon) N_p(\bm x | \bm \mu_k, \bm \Sigma_k) + \epsilon N_p(\bm x | \bm \mu_k, \eta \bm \Sigma_k),\label{eq:contaminated-pdf}
\end{align}
where $\epsilon \in [0, 1]$ is the probability that an observation is
contaminated (i.e., drawn from a distribution with larger variance) and $\eta >
1$ scales the covariance matrix $\bm \Sigma_k$ to increase the extremity of
outliers. For $\epsilon = 0$, we have the benchmark block-diagonal simulation
design from \cite{Guo:2007te}. As $\epsilon$ is increased, the average number of
outliers is increased. In our simulation, we set $\eta = 100$ and considered the
values of $\epsilon = 0$, $0.05$, $\ldots$, $0.50$.

We generated $K=3$ populations from \eqref{eq:contaminated-pdf} with $\bm
\Sigma_k$ given in \eqref{eq:block-sigma}. We set the mean vector of class 1 to
$\bm \mu_1 = \bm 0_p$. Next, comparable to \cite{Guo:2007te}, the first 100
features of $\bm \mu_2$ were 1/2, while the rest were set to 0, i.e., $\bm \mu_2
= (\underbrace{1/2, \ldots, 1/2}_{100}, \underbrace{0, \ldots, 0}_{p -
  100})$. For simplicity, we defined $\bm \mu_3 = -\bm \mu_2$. The three
populations differed in their mean vectors in the first 100 features
corresponding to the first block. No difference in the means occurred in the
remaining blocks.

From each of the $K=3$ populations, we sampled 25 training observations ($n_k =
25$ for all $k$) and 10,000 test observations. After training each classifier on
the training data, we classified the test data sets and computed the proportion
of mislabeled test observations to estimate the classification error rate for
each classifier. Repeating this process 500 times, we computed the average of
the error-rate estimates for each classifier. We allowed the number of features
to vary from $p = 100$ to $p = 500$ in increments of 100 to examine the
classification accuracy as the feature dimension increased while maintaining a
small sample size. \cite{Guo:2007te} originally considered $\rho_k = 0.9$ for
all $k$. To explore the more realistic assumption of unequal covariance
matrices, we instead put $\rho_1 = 0.1$, $\rho_2 = 0.5$, and $\rho_3 = 0.9$

\subsubsection{Simulation Results}

In Figure \ref{fig:sim-results}, we observe each classifier's average
classification error rates for the values of $\epsilon$ and $p$. Unsurprisingly,
the average error rate increased for each classifier as the contamination
probability $\epsilon$ increased regardless of the value of $p$. Sensitivity to
the presence of outliers was most apparent for the for the Pang, Tong, and
Witten classifiers. For smaller dimensions, the random-forest and \emph{HDRDA}
classifiers tended to outperform the remaining classifiers with the random
forest performing best. As the feature dimension increased with $p \ge 300$,
both \emph{HDRDA} classifiers outperformed all other classifiers, suggesting
that their inherent dimension reduction better captured the classificatory
information in the small samples even in the presence of outliers.

\begin{figure}
<<sim_results, fig.width=8, fig.height=8, dpi=200>>=
p <- ggplot(results_contaminated, aes(x=contamination_prob, y=Error))
p <- p + geom_line(aes(color=Classifier, linetype=Classifier), size=0.65)
p <- p + facet_wrap(~ p, labeller = "label_both")
p <- p + xlab("Contamination Probability") + ylab("Average Error Rate")
p + theme_bw()
@
\caption{Average classification error rates as a function of the contamination
  probability $\epsilon$. Approximate standard errors were no greater than
  \Sexpr{max_sim_se}.}
\label{fig:sim-results}
\end{figure}

The Pang, Tong, and Witten methods yielded practically the same and consistently
the worst error rates when outliers were present with $\epsilon > 0$, suggesting
that these classifiers were sensitive to outliers. Notice, for example, that
when $p=400$, the error rates of the Pang, Tong, and Witten classifiers
increased dramatically from approximately 19\% when no outliers were present to
approximately 43\% when $\epsilon=0.05$. The sharp increase in average error
rates for these three classifiers continued as $\epsilon$ increased
further. Guo's method always outperformed Pang, Witten, and Tong, but once
outliers were present, the Guo classifier's average error rate was competitive
with neither the \emph{HDRDA} classifiers nor the random forest.

\begin{figure}
<<sim_results_by_p, fig.width=8, fig.height=8, dpi=200>>=
p2 <- results_contaminated %>%
  mutate(p=as.numeric(as.character(p)),
         contamination_prob=round(contamination_prob, digits=2)) %>%
    filter(contamination_prob %in% c(0.0, 0.1, 0.2, 0.3, 0.4, 0.5)) %>%
    rename(`Contamination Probability`=contamination_prob) %>%
      ggplot(., aes(x=p, y=Error))
p2 <- p2 + geom_line(aes(color=Classifier, linetype=Classifier), size=0.65)
p2 <- p2 + facet_wrap(~ `Contamination Probability`, labeller="label_both")
p2 <- p2 + xlab("Number of Features") + ylab("Average Error Rate")
p2 + theme_bw()
@
\caption{Average classification error rates as a function of the number of
  features $p$. Approximate standard errors were no greater than
  \Sexpr{max_sim_se}.}
\label{fig:sim-results-by-p}
\end{figure}

In Figure \ref{fig:sim-results-by-p} we again examine the simulation results as
a function of the number of features $p$ for a subset of the values of
$\epsilon$. This set of plots allows us to investigate the effect of feature
dimensionality on classification performance. When no outliers were present
(i.e., $\epsilon=0$), the random-forest classifier was outperformed by all other
classifiers. Furthermore, the \emph{HDRDA} classifiers were superior in terms of
average error rate in this setting. As $p$ increased, an elevation in average
error rate was expected for all classifiers, but the increase was not observed
to be substantial.

For $\epsilon > 0$, we observed a different behavior in classification
performance. First, the Pang, Tong, and Witten methods along with the random
forest increased in average error rate as $p$ increased. Contrarily, the
performance of the \emph{HDRDA} and Guo classifiers was hardly affected by $p$.
Also, as discussed above, the \emph{HDRDA} classifiers were superior to all
other classifiers for larger values of $p$ with only the random forest
outperforming them for smaller feature dimensions.

\subsection{Application to Gene Expression Data}

We compared the \emph{HDRDA} classifer to the five competing classifiers on four
benchmark gene-expression microarray data sets. We evaluated the classification
performance of each classifier by randomly partitioning the data set under
consideration such that $2/3$ of the observations were allocated as training
data and the remaining $1/3$ of the observations were allocated as a test data
set. After training each classifier on the training data, we classified the test
data sets and computed the proportion of mislabeled test observations to
estimate the classification error rate for each classifier. Repeating this
process 250 times, we computed the average of the error-rate estimates for each
classifier. Here, we provide a concise description of each high-dimensional
data set examined in our classification study.

\subsubsection{\cite{Burczynski:2006ik} Data Set}

\cite{Burczynski:2006ik} acquired peripheral blood mononuclear cells (PBMC)
through hybridization to microarrays from 127 individuals resulting in 22,283
sequences. Of the 127 individuals, 42 were healthy, 59 had Crohn's disease (CD),
and 26 had ulcerative colitis (UC). The goal of \cite{Burczynski:2006ik} was to
improve accuracy in the discrimination of Inflammatory Bowel Disease (IBD) using
the PBMC-based gene expression signature of a patient.

\subsubsection{\cite{Nakayama:2007fl} Data Set}

\cite{Nakayama:2007fl} acquired 105 gene-expression samples of 10 types of soft
tissue tumors through an oligonucleotide microarray, including 16 samples of
synovial sarcoma (SS), 19 samples of myxoid/round cell liposarcoma (MLS), 3
samples of lipoma, 3 samples of well-differentiated liposarcoma (WDLS), 15
samples of dedifferentiated liposarcoma (DDLS), 15 samples of myxofibrosarcoma
(MFS), 6 samples of leiomyosarcoma (LMS), 3 samples of malignant nerve sheathe
tumor (MPNST), 4 samples of fibrosarcoma (FS), and 21 samples of malignant
fibrous histiocytoma (MFH). \cite{Nakayama:2007fl} determined from their data
that these 10 types fell into 4 broader groups: (1) SS; (2) MLS; (3) Lipoma,
WDLS, and part of DDLS; (4) Spindle cell and pleomorophic sarcomas including
DDLS, MFS, LMS, MPNST, FS, and MFH. Following \cite{Witten:2011kc}, we restrict
our analysis to the five tumor types having at least 15 observations.

\subsubsection{\cite{Shipp:2002ka} Data Set}

Approximately 30-40\% of adult non-Hodgkin lymphomas are diffuse large B-cell
lymphomas (DLBCLs) \cite{Shipp:2002ka}.  However, only a small proportion of
DLBCL patients are cured with modern chemotherapeutic regimens.  Several models
have been proposed, such as the International Prognostic Index (IPI), to
determine a patient's curability.  These models rely on clinical covariates,
such as age, to determine if the patient can be cured and are often
ineffective. \cite{Shipp:2002ka} have argued that more effective means are
desired to determine a patient's curability. The authors measured 6,817 gene
expression levels from 58 DLBCL patient samples with customized cDNA
(lymphochip) microarrays to investigate the curability of patients treated with
cyclophosphamide, adriamycin, vincristine, and prednisone (CHOP)-based
chemotherapy. Among the 58 DLBCL patient samples, 32 are from cured patients
while 26 are from patients with fatal or refractory disease.

\subsubsection{\cite{Singh:2002fh} Data Set}

\cite{Singh:2002fh} have examined 235 radical prostatectomy specimens from
surgery patients between 1995 and 1997. The authors used oligonucleotide
microarrays containing probes for approximately 12,600 genes and expressed
sequence tags. They have reported that 102 of the radical prostatectomy
specimens are of high quality: 52 prostate tumor samples and 50 non-tumor
prostate samples.

\subsubsection{Classification Results}

Similar to \cite{Witten:2011kc}, we report the average test error rates obtained
over 250 random training-test partitions in Table \ref{tab:microarray-results}
along with standard deviations of the test error rates in parentheses. The
\emph{HDRDA} and Guo classifiers were superior in classification performance for
the majority of the simulations with the popular Guo classifier providing
superior classification performance for the Burczysnki data set. Both
\emph{HDRDA} methods yielded comparable results.

<<classification_results, results='asis'>>=
table_caption <- 'The average of the test error rates obtained on gene-expression data sets over 250 random training-test partitions. Standard deviations of the test error rates are given in the parentheses.'

xtable_out <- xtable(results_microarray,
                     caption=table_caption,
                     label='tab:microarray-results')
print(xtable_out, include.rownames=FALSE)
@

The average error-rate estimates for the Pang, Tong, and Witten classifiers were
comparable across all data sets. Furthermore, the average error rates for the
Pang and Tong classifiers were approximately equal throughout our study,
suggesting that the mean and variance estimators used in lieu of the MLEs
provided little improvement to classification accuracies. The random-forest
classifier was competitive only on the Singh data set. This result challenges
the claim of \cite{FernandezDelgado:2014ul} that the random forest is
typically superior. More studies should be performed to validate this statement
in the small-sample, high-dimensional setting.

Finally, the average test error rates for the Witten classifier were the largest
across all data sets. Given that the standard deviations were relatively large,
we hesitate to generalize claims regarding the Witten classifier being the
uniformly worst performer. However, the variability in its classification
accuracy across multiple random partitions of each data set were large enough to
rethink applying it in practice.

\section{Discussion}
\label{sec:discussion}

We have compared our proposed \emph{HDRDA} classifier to the random-forest and
Witten classifiers as well as the Pang, Tong, and Guo classifiers, each of which
assumes independence of the features conditional on class membership. Despite
the rapid computational performance of diagonal classifiers and the reduction in
the number of parameters to estimate, \cite{Mai:2012bf} and \cite{Fan:2012iq}
have noted that diagonal classifiers can often yield inferior classification
performance to other classification methods. We have shown that the \emph{HDRDA}
classifier often yields superior classification accuracy to the diagonal
classifiers considered when applied to small-sample, high-dimensional data sets,
confirming the assertions of \cite{Mai:2012bf} and \cite{Fan:2012iq}.
Furthermore, we have demonstrated that \emph{HDRDA} classifiers are more robust
to the presence of outliers than the diagonal classifiers.

We also considered the popular penalized linear discriminant analysis from
\cite{Witten:2011kc} because it was specifically designed for high-dimensional
gene-expression data. We had expected its classification performance to be
competitive within our classification study and perhaps superior. Contrarily,
its underperformance in our empirical studies show that it is sensitive to
outliers and unable to achieve comparable results with other classifiers
designed for small-sample, high-dimensional data. Also, despite the claims of
\cite{FernandezDelgado:2014ul} that the random forest classifier is typically
superior to other classifiers, we observed that the random-forest classifier was
indeed competitive but was typically outperformed by classifiers developed for
small-sample, high-dimensional data.

The \emph{HDRDA} implementation of \eqref{eq:hdrda-decomposed} in the {\tt
  sparsediscrim} R package was comparable to that of the diagonal classifiers in
terms of computational runtime. Given the \emph{RDA} classifier from
\cite{Friedman:1989tm} has been shown to have excellent performance in the
high-dimensional setting \citep{Webb:2011vu} but is limited by its
computationally intense model selection procedure, our work enables the
\emph{RDA} classifier to be employed with small-sample, high-dimensional data in
practice. This results is reassuring because the \emph{RDA} classifier remains
widely popular in the literature. In fact, variants of the RDA classifier have
been applied to microarray data
\citep{Ching:2012fu,Li:2012ev,Tai:2007bk,Guo:2007te}, facial recognition
\citep{Ching:2012fu,Zhang:2010va,Dai:2007vd,Lu:2005hq,Pima:2004vw,Lu:2003we},
handwritten digit recognition \citep{Bouveyron:2007gx}, remote sensing
\citep{Tadjudin:1999fk}, seismic detection \citep{Anderson:2002kg}, and chemical
spectra \citep*{Wu:1996us,Aeberhard:1993fp}.

The dimension reduction that we have employed in this paper has reduced the
dimension to rank$(\widehat{\bm \Sigma}) = q$. An interesting extension of
our work would reduce the dimension $q$ further to a lower dimension $q_L < q$
using some criterion proposed, perhaps similar to that employed in principal
components analysis. While it is not clear whether classification performance
would improve via such a method, the efficiency of the training and model
selection would certainly improve. Moreover, if $q_L = 2$ or $3$ was selected,
the potential for low-dimensional graphical displays of high-dimensional could
be accomplished.

\bibliographystyle{plainnat}
\bibliography{rda}

\end{document}