### 2016-01-30 reports/paper-hdrda.Rnw
Applied edits from Mrs. Young #23

\documentclass{article}

\usepackage{graphicx, amsmath, amssymb, bm, mathtools, amsthm, natbib, setspace, algorithm2e}

\bibpunct{(}{)}{;}{a}{,}{,}

\newcommand{\xbar}{\bar{\bm x}}
\newcommand{\tr}{\text{tr}}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\newtheorem{cor}{Corollary}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{thm}{Theorem}

\pagestyle{plain}
%----------------Page dimensions ----------------
\oddsidemargin 0.0in
\evensidemargin 0.0in
\topmargin -0.75in
\leftmargin 0in
\headheight 0.0in
\headsep 0.5in
\footskip 0.5in
\footnotesep 0.0in
\textwidth 6.5in
\textheight 9.5in

\begin{document}

\title{High-Dimensional Regularized Discriminant Analysis}

\author{John A. Ramey, Caleb K. Stein, Phil D. Young, and Dean M. Young}

\maketitle

\doublespacing

\begin{abstract}

\citeauthor{Friedman:1989tm} proposed the popular regularized discriminant
analysis (RDA) classifier that utilizes a biased covariance-matrix estimator
that partially pools the sample covariance matrices from linear and quadratic
disriminant analysis and shrinks the resulting estimator towards a scaled
identity matrix. The RDA classifier's two tuning parameters are typically
estimated via a computationally burdensome cross-validation procedure that uses
a grid search. We formulate a new RDA-based classifier for the small-sample,
high-dimensional setting and then show that the classification decision rule is
equivalent to a classifier in a subspace having a much lower dimension. As a
result, the utilization of the dimension-reduction step yields a substantial
reduction in computation during model selection. Also, our parameterization
offers interpretability that was previously lacking with the RDA classifier. We
demonstrate that our proposed classifier is often superior to several recently
proposed sparse and regularized classifiers in terms of classification accuracy
with three artificial and six real high-dimensional data sets. Finally, we
provide an implementation of our proposed classifier in the {\tt sparsediscrim}
R package, which is available on CRAN.

\end{abstract}

<<knitr_setup, include=FALSE, echo=FALSE, cache=FALSE>>=
# Round digits throughout document including \Sexpr statements.
# See Yihui's response here:
# http://stackoverflow.com/questions/11062497/how-to-avoid-using-round-in-every-sexpr
options(digits=3)

opts_chunk$set(fig.align='default', dev='png', message=FALSE, warning=FALSE,
               cache=TRUE, echo=FALSE, autodep=TRUE, fig.width=12,
               fig.height=12, out.width="\\linewidth")
@

<<setup>>=
setwd("..")
library(ProjectTemplate)
load.project()
@

<<munge_contaminated_sims>>=
# Contaminated Normal Sims
results_contaminated <- lapply(results_block_contaminated[[1]], function(r) {
  with(r, c(error_rates, p=p, mu=mu, contamination_prob=contamination_prob))
})
results_contaminated <- do.call(rbind.data.frame, results_contaminated)
results_contaminated <- tbl_df(results_contaminated)

results_contaminated <- results_contaminated %>%
  group_by(p, contamination_prob, mu) %>%
  summarise_each(funs(mean)) %>%
  arrange(p, contamination_prob, mu) %>%
  ungroup

results_contaminated <- results_contaminated %>%
  gather(Classifier, Error, -p, -contamination_prob, -mu) %>%
  mutate(p=factor(p)) %>%
  filter(mu == 0.5)

# Remove underscore from classifier names
levels(results_contaminated$Classifier) <- gsub("_", " ",
                                                levels(results_contaminated$Classifier))

# Maximum Standard Error for Contaminated Sims
max_sim_se <- with(results_contaminated, max(sqrt(Error * (1 - Error) / 500)))
max_sim_se <- round(max_sim_se, digits=3)
@

<<munge_microarray_results>>=
# Microarray Classification Results
results_microarray <- lapply(results[[1]], function(r) {
  with(r, c(error_rates, `data_set`=data_set))
})
results_microarray <- do.call(rbind.data.frame, results_microarray)
results_microarray <- tbl_df(results_microarray)

avg_error_microarray <- results_microarray %>%
    group_by(data_set) %>%
    summarise_each(funs(mean)) %>%
    gather(Classifier, Avg_Error, -data_set)

sd_error_microarray <- results_microarray %>%
  group_by(data_set) %>%
  summarise_each(funs(sd)) %>%
  gather(Classifier, SD_Error, -data_set)

# Microarray classification results are presented in a table
# Rows = classifier, columns = data set
# Cells = Error (SD) rounded to 3 digits
# The classifier with the minimum average error rate is in bold for each data set.

bold_latex <- function(x) {
  paste0('\\textbf{', x, '}')
}

results_microarray <- inner_join(avg_error_microarray, sd_error_microarray,
                                 by=c('data_set', 'Classifier')) %>%
  mutate(Table_Cells=sprintf('%.3f', Avg_Error),
         data_set=stringr::str_to_title(data_set),
         Classifier=gsub("_", " ", Classifier)) %>%
  group_by(data_set) %>%
  mutate(Table_Cells=ifelse(Avg_Error == min(Avg_Error), bold_latex(Table_Cells), Table_Cells),
         Table_Cells=sprintf('%s (%.3f)', Table_Cells, SD_Error)) %>%
  select(-Avg_Error, -SD_Error) %>%
  spread(data_set, Table_Cells)
@

\section{Introduction}

As machine-learning algorithms are increasingly utilized to automate data-driven
decisions, researchers continue to develop a wealth of \emph{classifiers}
employed in a variety of disciplines, such as medical research, facial
recognition, remote sensing, and chemical analysis
\citep*{Ching:2012fu,Li:2012ev,Zhang:2010va}. Despite the advances in
classification to facilitate technological advancements, well-established
classifiers often degrade in accuracy when applied to modern, high-dimensional
data sets if the number of features $p$ exceeds the total sample size $N$, due
to the \emph{curse of dimensionality} \citep{Bellman:1961tn}. For instance,
linear discriminant analysis (\emph{LDA}) and quadratic discriminant analysis
(\emph{QDA}) are popular because of their simplicity, speed of computation, and
interpretability. However, both methods are incalculable when $p > N$ because
the class sample covariance matrices are singular \citep*{Bouveyron:2007gx}.

As a result, researchers have emphasized three standard approaches to modifying
the \emph{LDA} and \emph{QDA} classifiers for application to high-dimensional
data: 1) covariance-matrix regularization, 2) dimension reduction, and 3)
parameter restrictions. Covariance-matrix regularization methods are often based
on the well-known ridge-regression approach of \cite{Hoerl:1970cd}.
Additionally, \cite{Ching:2012fu}, \cite{Zhang:2010va}, \cite{Ye:2009gd},
\cite{Ji:2008wp}, \cite*{Guo:2007te}, \cite*{Srivastava:2007ud},
\cite{Ye:2006vx}, and \cite{Ye:2005uu} have proposed regularization methods to
improve the estimation of the class covariance matrices by adjusting the
eigenvalues of the sample covariance matrices to ensure positive
definiteness. Alternatively, dimension reduction typically transforms the
original data in the high-dimensional space to a subspace having a much smaller
dimension such that redundant and noisy information is removed
\citep*{Fan:2012iq,Ye:2009gd,Bouveyron:2007gx,Ye:2006vx,Ye:2005hd,Antoniadis:2003bk}.
Finally, \cite*{Mai:2012bf}, \cite{Fan:2012iq}, \cite*{Merchante:2012vk},
\cite*{Clemmensen:2011kr}, \cite*{Huang:2010ju}, and \cite*{Dudoit:2002ev} have
proposed restrictions on the population covariance matrices to reduce the number
of parameters that must be estimated. For instance, \cite{Dudoit:2002ev} have
assumed that the features within each class are independent so that the
population covariance matrices are diagonal, resulting in positive-definite
sample covariance matrices. For excellent overviews of the regularization
problem in discriminant analysis and other techniques, see \cite{Murphy:2012uq}
and \cite*{Mkhadri:1997gy}.

Here, we present the high-dimensional \emph{RDA} (\emph{HDRDA}) classifier,
which is intended for the $p > N$ case and builds on the classic \emph{RDA}
classifier proposed by \cite{Friedman:1989tm}. In this paper, we study the
classification performance of the \emph{HDRDA} classifier on six real
high-dimensional data sets along with a simulation design that generalizes the
experiments initially conducted by \cite{Guo:2007te}. We demonstrate that the
\emph{HDRDA} classifier often attains superior classification accuracy to
several recent classifiers designed for small-sample, high-dimensional data from
\cite*{Tong:2012hw}, \cite{Witten:2011kc}, \cite*{Pang:2009ik}, and
\cite{Guo:2007te}. We also include as a benchmark the random forest from
\cite{Breiman:2001fb} because \cite*{FernandezDelgado:2014ul} have concluded
that the random forest is often superior to other classifiers in benchmark
studies. We show that our proposed classifier is competitive and often
outperforms the random forest in terms of classification accuracy in the
small-sample, high-dimensional setting.

Our proposed \emph{HDRDA} classifier employs a biased covariance-matrix
estimator that partially pools the individual sample covariance matrices from
the \emph{QDA} classifier with the pooled sample covariance matrix from the
\emph{LDA} classifier. We then shrink the resulting covariance-matrix estimator
towards a scaled identity matrix to ensure positive definiteness. By
reparameterizing the \emph{RDA} classifier similar to that of
\cite*{Hastie:2008dt} and \cite{Halbe:2007ea}, we show that the pooling
parameter in the \emph{HDRDA} classifier determines the contribution of each
training observation to the estimation of each class covariance matrix, enabling
interpretability that has been previously lacking with the \emph{RDA} classifier
\citep{Bensmail:1996fw}.

Next, we achieve a substantial reduction in dimension that greatly reduces the
computational effort during model selection when $p \gg N$ by applying reasoning
similar to that of \cite{Ye:2006tq}. We establish that the matrix operations
corresponding to the null space of the pooled sample covariance matrix are
redundant and can be discarded from the \emph{HDRDA} decision rule without loss
of classificatory information. Moreover, because the \emph{HDRDA} decision rule
is invariant to adjustments to the zero eigenvalues, the decision rule in the
original feature space is equivalent to a decision rule in a lower dimension,
such that matrix inverses and determinants of relatively small matrices can be
rapidly computed. Also, we provide an efficient algorithm along with pseudocode
to estimate the \emph{HDRDA} classifier's tuning parameters in a grid search via
cross-validation. Finally, we show that several shrinkage methods that are
special cases of the \emph{HDRDA} classifier have no effect on the zero
eigenvalues of the covariance-matrix estimators. Such techniques include work
from \cite{Srivastava:2007ww}, \cite{Rao:1971ul}, and other methods studied by
\cite{Ramey:2013ji} and \citet*{Xu:2009fl}.

The remainder of this paper is organized as follows. In Section
\ref{sec:preliminaries} we present the discrimination problem and necessary
notation to describe our contributions. In Section \ref{sec:rda} we present the
\emph{HDRDA} classifier along with its interpretation, properties, and
computationally efficient model-selection procedure. In Section \ref{sec:sims}
we describe our simulation studies of artificial and real data sets and describe
the experimental results. We conclude with a brief discussion in Section
\ref{sec:discussion}.

\section{Preliminaries}
\label{sec:preliminaries}

\subsection{Notation}

To facilitate our discussion of covariance-matrix regularization and regularized
discriminant analysis, we require the following notation. Let $\mathbb{R}_{a
  \times b}$ denote the matrix space of all $a \times b$ matrices over the real
field $\mathbb{R}$. Denote by $\bm I_m$ the $m \times m$ identity matrix, and
let $\bm 0_{m \times p}$ be the $m \times p$ matrix of zeros, such that $\bm
0_m$ is understood to denote $\bm 0_{m \times m}$. Define $\bm 1_m \in
\mathbb{R}_{m \times 1}$ as a vector of ones. Let $\bm A \oplus \bm B$ denote
the direct sum of $\bm A \in \mathbb{R}_{r \times r}$ and $\bm B \in
\mathbb{R}_{s \times s}$ \citep[Chapter 1]{Lutkepohl:1996uz}. Let $\bm A^{T}$,
$\bm A^+$, and $\mathcal{N}(\bm A)$ denote the transpose, the Moore-Penrose
pseudoinverse, and the null space of $\bm A \in \mathbb{R}_{m \times p}$,
respectively. Denote by $\mathbb{R}_{p \times p}^{>}$ the cone of real $p \times
p$ positive-definite matrices. Similarly, let $\mathbb{R}_{p \times p}^{\ge}$
denote the cone of real $p \times p$ positive-semidefinite matrices. Let
$V^{\perp}$ denote the orthogonal complement of a vector space $V \subset
\mathbb{R}_{p \times 1}$. For $c \in \mathbb{R}$, let $c^+ = 1/c$ if $c \ne 0$
and $0$ otherwise.

\subsection{Discriminant Analysis}

In discriminant analysis we wish to assign an unlabeled vector $\bm x \in
\mathbb{R}_{p \times 1}$ to one of $K$ unique, known classes by constructing a
classifier from $N$ training observations. Let $\bm x_i = (x_{i1}, \ldots,
x_{ip}) \in \mathbb{R}_{p \times 1}$ be the $i$th observation $(i = 1, \ldots,
N)$ with true, unique membership $y_i \in \{\omega_1, \ldots, \omega_K\}$.
Denote by $n_k$ the number of training observations realized from class $k$,
such that $\sum_{k=1}^K n_k = N$. We assume that $(\bm x_i, y_i)$ is a
realization from a mixture distribution $p(\bm x) = \sum_{k=1}^K p(\bm x |
\omega_k) p(\omega_k)$, where $p(\bm x | \omega_k)$ is the probability density
function (PDF) of the $k$th class and $p(\omega_k)$ is the prior probability of
class membership of the $k$th class. We further assume $p(\omega_k) =
p(\omega_l)$, $1 \le k, l \le K$, $k \ne l$.

The \emph{QDA} classifier is the optimal Bayesian decision rule with respect to
a $0-1$ loss function when $p(\bm x | \omega_k)$ is the PDF of the multivariate
normal distribution with known mean vectors $\bm \mu_k \in \mathbb{R}_{p \times 1}$
and known covariance matrices $\bm \Sigma_k \in \mathbb{R}_{p \times p}^{>}$,
$k = 1, 2, \ldots, K$. Because $\bm \mu_k$ and $\bm \Sigma_k$ are typically
unknown, we assign an unlabeled observation $\bm x$ to class $\omega_k$ with the
sample \emph{QDA} classifier
\begin{align}
	D_{QDA}(\bm x) = \argmin_{k} (\bm x - \xbar_k)^{T}\widehat{\bm
    \Sigma}_k^{-1}(\bm x - \xbar_k) + \log |\widehat{\bm \Sigma}_k|, \label{eq:qda}
\end{align}
where $\xbar_k$ and $\widehat{\bm \Sigma}_k$ are the maximum likelihood
estimators (MLEs) of $\bm \mu_k$ and $\bm \Sigma_k$, respectively. If we assume
further that $\bm \Sigma_k = \bm \Sigma$, $k = 1, \ldots, K$, then the pooled
sample covariance matrix $\widehat{\bm \Sigma}$ is substituted for $\widehat{
  \bm \Sigma}_k$ in \eqref{eq:qda}, where
\begin{align}
	\widehat{\bm \Sigma} = N^{-1} \sum_{k=1}^K n_k \widehat{\bm \Sigma}_k \label{eq:pooled-cov}
\end{align}
is the MLE for $\bm \Sigma$. Here, \eqref{eq:qda} reduces to the sample
\emph{LDA} classifier. We omit the log-determinant because it is constant across
the $K$ classes.

The smallest eigenvalues of $\widehat{\bm \Sigma}_k$ and the directions
associated with their eigenvectors can highly influence the classifier in
\eqref{eq:qda}. In fact, the eigenvalues of $\widehat{ \bm \Sigma}_k$ are
well-known to be biased if $p \ge n_k$ such that the smallest eigenvalues are
underestimated \citep{Seber:2004uh}. Moreover, if $p > n_k$, then
rank$(\widehat{\bm \Sigma}_k) \le n_k$, which implies that at least $p - n_k$
eigenvalues of $\widehat{\bm \Sigma}_k$ are zero. Furthermore, although more
feature information is available to discriminate among the $K$ classes, if $p >
n_k$, \eqref{eq:qda} is incalculable because $\widehat{\bm \Sigma}_k^{-1}$ does
not exist.

Several regularization methods, such as the methods considered by
\cite{Xu:2009fl}, \cite{Guo:2007te}, and \cite{Mkhadri:1995jp}, have been
proposed in the literature to adjust the eigenvalues of $\widehat{\bm \Sigma}_k$
so that \eqref{eq:qda} is calculable and provides reduced variability for
$\widehat{\bm \Sigma}_k^{-1}$. A common form of the covariance-matrix
regularization applies a shrinkage factor $\gamma > 0$, so that
\begin{align}
	\widehat{\bm \Sigma}_k(\gamma) = \widehat{\bm \Sigma}_k + \gamma \bm I_p, \label{eq:ridge-estimator}
\end{align}
similar to a method employed in ridge regression \citep{Hoerl:1970cd}. Equation
\eqref{eq:ridge-estimator} effectively \emph{shrinks} the sample covariance
matrix $\widehat{\bm \Sigma}_k$ toward $\bm I_p$, thereby increasing the
eigenvalues of $\widehat{\bm \Sigma}_k$ by $\gamma$. Specifically, the zero
eigenvalues are replaced with $\gamma$, so that \eqref{eq:ridge-estimator} is
positive definite. For additional covariance-matrix regularization methods, see
\cite{Ramey:2013ji}, \cite{Xu:2009fl}, and \cite{Ye:2009gd}.

\section{High-Dimensional Regularized Discriminant Analysis}
\label{sec:rda}

Here, we define the \emph{HDRDA} classifier by first formulating the
covariance-matrix estimator $\widehat{\bm \Sigma}_k(\lambda)$ and demonstrating
its clear interpretation as a linear combination of the crossproducts of the
training observations centered by their respective class sample means. We define
the convex combination
\begin{align}
  \widehat{\bm \Sigma}_k(\lambda) \coloneqq (1 - \lambda) \widehat{\bm \Sigma}_k + \lambda
  \widehat{\bm \Sigma}, \quad k = 1, \ldots, K,\label{eq:sig-lambda-alternative}
\end{align}
where $\lambda \in [0, 1]$ is the \emph{pooling} parameter. By rewriting
\eqref{eq:sig-lambda-alternative} in terms of the observations $\bm x_i$, $i =
1, \ldots, N$, each centered by its class sample mean, we attain a clear
interpretation of $\widehat{\bm \Sigma}_k(\lambda)$. That is,
\begin{align}
	\widehat{\bm \Sigma}_k(\lambda)
  &= \left( 1 - \lambda + \frac{\lambda n_k}{N} \right) \widehat{\bm \Sigma}_k +  \frac{\lambda}{N} \sum_{\substack{k' = 1\\k' \ne k}}^K n_{k'} \widehat{\bm \Sigma}_{k'} \nonumber \\
	&= \left( \frac{1 - \lambda}{n_k} + \frac{\lambda}{N} \right)\sum_{i=1}^N
  I(y_i = k) \bm x_i \bm x_i^{T} +  \frac{\lambda}{N} \sum_{i=1}^N I(y_i \ne k)
  \bm x_i \bm x_i^{T} \nonumber \\
	&= \sum_{i=1}^N c_{ik}(\lambda) \bm x_i \bm x_i^{T},\label{eq:sig-lambda-alternative2}
\end{align}
where $c_{ik}(\lambda) = \lambda N^{-1} + (1 - \lambda)n_k^{-1}I(y_i = k)$.
From \eqref{eq:sig-lambda-alternative2}, we see that $\lambda$ weights the
contribution of each of the $N$ observations in estimating $\bm \Sigma_k$ from
all $K$ classes rather than using only the $n_k$ observations from a single
class. As a result, we can interpret \eqref{eq:sig-lambda-alternative2} as a
covariance-matrix estimator that borrows from $\widehat{\bm \Sigma}$ in
\eqref{eq:pooled-cov} to estimate $ \bm \Sigma_k$.

In Figure \ref{fig:hdrda-contours} we plot the contours of five multivariate
normal populations for $\lambda = 0$ with unequal covariance matrices. As
$\lambda$ approaches 1, the contours become more similar, resulting in identical
contours for $\lambda = 1$. Below, we show that the pooling operation is
advantageous in increasing the rank of each $\widehat{\bm \Sigma}_k(\lambda)$
from rank$(\widehat{\bm \Sigma}_k)$ to rank$(\widehat{\bm \Sigma})$ for $0 <
\lambda \le 1$. Notice that if $\lambda = 0$, then the observations from the
remaining $K - 1$ classes do not contribute to the estimation of $\bm \Sigma_k$,
corresponding to $\widehat{\bm \Sigma}_k$. Furthermore, if $\lambda = 1$, the
weights $c_{ik}(\lambda)$ in \eqref{eq:sig-lambda-alternative2} reduce to $1/N$,
corresponding to $\widehat{\bm \Sigma}$. For brevity, when $\lambda = 1$, we
define $\bm X = [\sqrt{c_{1k}(1)} \bm x_1^{T}, \ldots, \sqrt{c_{Nk}(1)} \bm
  x_N^{T}]^{T}$ such that $\widehat{\bm \Sigma} = N^{-1} \bm X^{T} \bm
X$. Similarly, for $\lambda = 0$, we define $\bm X_k = [\sqrt{c_{1k}(0)} \bm
  x_1^{T}, \ldots, \sqrt{c_{Nk}(0)} \bm x_N^{T}]^{T}$ such that $\widehat{\bm
  \Sigma}_k = n_k^{-1} \bm X_k^{T} \bm X_k$.

\[ \left[\text{Insert Figure \ref{fig:hdrda-contours} approximately here }\right] \]

As we have discussed above, several eigenvalue adjustment methods have been
proposed that increase eigenvalues (approximately) equal to 0. To further
improve the estimation of $\bm \Sigma_k$ and to stabilize the estimator's
inverse, we define the eigenvalue adjustment of
\eqref{eq:sig-lambda-alternative} as
\begin{align}
	\tilde{\bm \Sigma}_k \coloneqq \alpha_k \widehat{\bm \Sigma}_k(\lambda) + \gamma \bm I_p,\label{eq:hdrda-cov}
\end{align}
where $\alpha_k \ge 0$ and $\gamma \ge 0$ is an eigenvalue-shrinkage
constant. Thus, the \emph{pooling} parameter $\lambda$ controls the amount of
estimation information borrowed from $\widehat{\bm \Sigma}$ to estimate $\bm
\Sigma_k$, and the \emph{shrinkage} parameter $\gamma$ determines the degree of
eigenvalue shrinkage. The choice of $\alpha_k$ allows for a flexible formulation
of covariance-matrix estimators. For instance, if $\alpha_k = 1$, $k = 1,
\ldots, K$, then \eqref{eq:hdrda-cov} resembles \eqref{eq:ridge-estimator}.
Similarly, if $\alpha_k = 1 - \gamma$, then \eqref{eq:hdrda-cov} has a form
comparable to the \emph{RDA} classifier from \cite{Friedman:1989tm}.
Substituting \eqref{eq:hdrda-cov} into \eqref{eq:qda}, we define the
\emph{HDRDA} classifier
\begin{align}
	D_{HDRDA}(\bm x) = \argmin_{k}  (\bm x - \xbar_k)^{T}\tilde{\bm \Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm \Sigma}_k|. \label{eq:hdrda}
\end{align}
For $\gamma > 0$, $\tilde{\bm \Sigma}_k$ is nonsingular such that $\tilde{\bm
  \Sigma}_k^{-1}$ can be substituted for $\tilde{\bm \Sigma}_k^{+}$ in
\eqref{eq:hdrda}. If $\gamma = 0$, we explicitly set $|\tilde{\bm \Sigma}_k|$
equal to the product of the positive eigenvalues of $\tilde{\bm
  \Sigma}_k$. Following \cite{Friedman:1989tm}, we select $\lambda$ and $\gamma$
from a grid of candidate models via cross-validation. We provide an
implementation of \eqref{eq:hdrda} in the {\tt hdrda} function contained in the
{\tt sparsediscrim} R package, which is available on CRAN.

The choice of $\alpha_k$ in \eqref{eq:hdrda-cov} is one of convenience and
allows the flexibility of various covariance-matrix estimators proposed in the
literature. In practice, we generally are not interested in estimating
$\alpha_k$ because the estimation of $K$ additional tuning parameters via
cross-validation is counterproductive to our goal of improving the \emph{RDA}
classifier's computational efficiency. By setting $\alpha_k$ equal to an
appropriate constant, the \emph{HDRDA} covariance-matrix estimator includes or
resembles a large family of estimators. Notice that if $\alpha_k = 1$ and
$\lambda = 1$, \eqref{eq:hdrda-cov} is equivalent to the standard ridge-like
covariance-matrix estimator in \eqref{eq:ridge-estimator}, and various
estimators proposed in the literature can be obtained when one selects $\gamma$
accordingly. For instance, with $\gamma = \tr\{ \widehat{\bm \Sigma} \} /
\min(N, p)$, we obtain the estimator from \cite{Srivastava:2007ww}.

Perhaps the most well-known of the ridge-like estimators is from
\cite{Friedman:1989tm}, who proposed regularized discriminant analysis (RDA) by
incorporating a weighted average of the sample covariance matrix $\widehat{ \bm
  \Sigma}_k$ and the pooled sample covariance matrix $\widehat{\bm \Sigma}$ to
estimate $\bm \Sigma_k$ with
\begin{align}
    \widehat{\bm \Sigma}_k(\lambda) = n_k^{-1}(\lambda) \bm S_k(\lambda) \quad
    (k =
      1, \ldots, K),\label{eq:sig-lambda}
      \end{align}
where $\lambda \in [0, 1]$, $n_k(\lambda) = (1 - \lambda) n_k + \lambda N$, $
\bm S_k = n_k \widehat{\bm \Sigma}_k$, $\bm S = \sum_{k=1}^K \bm S_k$, and $ \bm
S_k(\lambda) = (1 - \lambda) \bm S_k + \lambda \bm S$. If $\lambda = 0$,
\eqref{eq:sig-lambda} is equal to $\widehat{\bm \Sigma}_k$ in \eqref{eq:qda}.
Alternatively, if $\lambda = 1$, then \eqref{eq:sig-lambda} reduces to
\eqref{eq:pooled-cov}. To further improve the estimation of $\bm \Sigma_k$ and
to stabilize the inverse of \eqref{eq:sig-lambda}, \cite{Friedman:1989tm}
proposed the biased covariance-matrix estimator
\begin{align}
          \widehat{\bm \Sigma}_k(\lambda, \gamma) = (1 - \gamma) \widehat{\bm
            \Sigma}_k(\lambda) + \gamma \frac{\tr\{\widehat{\bm
              \Sigma}_k(\lambda)\}}{p} \bm I_p,\label{eq:sig-rda}
          \end{align}
where $\gamma \in [0, 1]$ is a regularization parameter that controls the
shrinkage of \eqref{eq:sig-rda} towards $\bm I_p$ and is weighted by the average
of the eigenvalues of \eqref{eq:sig-lambda}. Finally, \cite{Friedman:1989tm}
established the RDA classifier by substituting \eqref{eq:sig-rda} into
\eqref{eq:qda}.

Our proposed \emph{HDRDA} classifier resembles \eqref{eq:sig-rda} but yields an
interpretation of the pooling parameter $\lambda$ as discussed above. Moreover,
as we discuss below, our formulation of the \emph{HDRDA} classifier lends itself
to rapid estimation in high-dimensional classification.

\section{Properties of the HDRDA Classifier}
\label{sec:hdrda-properties}

Next, we establish properties of the covariance-matrix estimator and the
decision rule employed in the \emph{HDRDA} classifier. By doing so, we
demonstrate that \eqref{eq:hdrda} lends itself to a more efficient
calculation. We decompose \eqref{eq:hdrda} into a sum of two components, where
the first summand consists of matrix operations applied to low-dimensional
matrices and the second summand corresponds to the null space of $\widehat{\bm
  \Sigma}$ in \eqref{eq:pooled-cov}. We show that the matrix operations
performed on the null space of $\widehat{\bm \Sigma}$ yield constant quadratic
forms across all classes and can be omitted. For $p \gg N$, the constant
component involves determinants and inverses of high-dimensional matrices, and
by ignoring these calculations, we achieve a substantial reduction in
computational costs. Furthermore, a byproduct is that adjustments to the
associated eigenvalues have no effect on \eqref{eq:hdrda}. Lastly, we utilize
the singular value decomposition to efficiently calculate the eigenvalue
decomposition of $\widehat{\bm \Sigma}$, further reducing the computational
costs of the \emph{HDRDA} classifier.

First, we require the following relationship regarding the null spaces of
$\widehat{\bm \Sigma}_k(\lambda)$, $\widehat{\bm \Sigma}$, and $\widehat{ \bm
  \Sigma}_k$.

\begin{lemma}\label{lemma:null-spaces}
Let $\widehat{\bm \Sigma}_k$ and $\widehat{\bm \Sigma}$ be the MLEs of $\bm
\Sigma_k$ and $\bm \Sigma$, respectively. Let $\widehat{\bm \Sigma}_k(\lambda)$
be defined as in \eqref{eq:sig-lambda-alternative}. Then,
$\mathcal{N}\{\widehat{\bm \Sigma}_k(\lambda)\} \subset \mathcal{N}(\widehat{
  \bm \Sigma}) \subset \mathcal{N}(\widehat{\bm \Sigma}_k)$, $k = 1, \ldots, K$.
\end{lemma}
\begin{proof}
Let $\bm z \in \mathcal{N}\{\widehat{\bm \Sigma}_k(\lambda)\}$ for some $k = 1,
\ldots, K$. Hence, $0 = \bm z^{T} \widehat{\bm \Sigma}_k(\lambda) \bm z = (1 -
\lambda) \bm z^{T} \widehat{\bm \Sigma}_k \bm z + \lambda \bm z^{T} \widehat{
  \bm \Sigma} \bm z$. Because $\widehat{\bm \Sigma}_k, \widehat{\bm \Sigma}\in
\mathbb{R}_{p \times p}^{\ge}$, we have $\bm z \in \mathcal{N}(\widehat{
  \bm \Sigma})$ and $\bm z \in \mathcal{N}(\widehat{\bm \Sigma}_k)$. In particular,
we have that $\mathcal{N}\{\widehat{\bm \Sigma}_k(\lambda)\} \subset
\mathcal{N}(\widehat{\bm \Sigma})$. Now, suppose $\bm z \in
\mathcal{N}(\widehat{\bm \Sigma})$. Similarly, we have that $0 = \bm z^{T}
\widehat{\bm \Sigma} \bm z = N^{-1} \sum_{k = 1}^K n_k \bm z^{T} \widehat{
  \bm \Sigma}_k \bm z$, which implies that $\bm z \in \mathcal{N}(\widehat{
  \bm \Sigma}_k)$ because $\widehat{\bm \Sigma}_k \in \mathbb{R}_{p \times
  p}^{\ge}$. Therefore, $\mathcal{N}(\widehat{\bm \Sigma}) \subset
\mathcal{N}(\widehat{\bm \Sigma}_k)$.
\end{proof}

In Lemma \ref{lemma:rda-tilde-Sigma_k} below, we derive an alternative
expression for $\tilde{\bm \Sigma}_k$ in terms of the matrix of eigenvectors of
$\widehat{\bm \Sigma}$. Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U^{T}$ be the
eigendecomposition of $\widehat{\bm \Sigma}$ such that $\bm D \in \mathbb{R}_{p
  \times p}^{\ge}$ is the diagonal matrix of eigenvalues of $\widehat{\bm \Sigma}$
with $\bm D = \bm D_q \oplus \bm 0_{p-q}$, $\bm D_q \in \mathbb{R}_{q \times
  q}^{>}$ is the diagonal matrix consisting of the positive eigenvalues of
$\widehat{\bm \Sigma}$, the columns of $\bm U \in \mathbb{R}_{p \times p}$ are
the corresponding orthonormal eigenvectors of $\widehat{\bm \Sigma}$, and
rank$(\widehat{\bm \Sigma}) = q$. Then, we partition $\bm U = (\bm U_1,
\bm U_2)$ such that $\bm U_1 \in \mathbb{R}_{p \times q}$ and $\bm U_2 \in
\mathbb{R}_{p \times (p - q)}$.

\begin{lemma}\label{lemma:rda-tilde-Sigma_k}
Let $\widehat{\bm \Sigma} = \bm U \bm D \bm U^{T}$ be the eigendecomposition of
$\widehat{\bm \Sigma}$ as above, and suppose that rank$(\widehat{\bm \Sigma}) =
q \le p$. Then, we have
\begin{align}
	\tilde{\bm \Sigma}_k &= \bm U(\bm W_k \oplus \gamma \bm I_{p-q})\bm U^{T}, \quad
  k = 1, \ldots, K,\label{eq:rda-matrix}
\intertext{where}
\bm W_k &= \alpha_k \{(1 - \lambda) \bm U_1^{T} \widehat{\bm \Sigma}_k \bm U_1 + \lambda \bm D_q\} + \gamma \bm I_{q}.\label{eq:Wk}
\end{align}
\end{lemma}
\begin{proof}
From Lemma \ref{lemma:null-spaces}, the columns of $\bm U_2$ span the null space
of $\widehat{\bm \Sigma}_k$, which implies that $\widehat{\bm \Sigma}_k \bm U_2
= \bm 0_{p \times (p - q)}$. Hence, $\bm U^{T} \widehat{\bm \Sigma}_k \bm U =
\bm U_1^{T} \widehat{\bm \Sigma}_k \bm U_1 \oplus \bm 0_{p-q}$, $k = 1, \ldots,
K$. Thus, $\bm U^{T} \tilde{\bm \Sigma}_k \bm U = \alpha_k \{(1 - \lambda) \bm
U^{T} \widehat{\bm \Sigma}_k \bm U + \lambda \bm D\} + \gamma \bm I_p$, and
\eqref{eq:rda-matrix} holds because $\bm U$ is orthogonal.
\end{proof}

As an immediate consequence of Lemma \ref{lemma:rda-tilde-Sigma_k}, we have the
following corollary.

\begin{cor}
Let $\widehat{\bm \Sigma}_k(\lambda)$ be defined as in
\eqref{eq:sig-lambda-alternative}. Then, for $\lambda \in (0, 1]$,
  rank$\{\widehat{\bm \Sigma}_k(\lambda)\} = q$, $k = 1, \ldots, K$.
\end{cor}
\begin{proof}
The proof follows when we set $\gamma = 0$ in Lemma
\ref{lemma:rda-tilde-Sigma_k}.
\end{proof}

Thus, by incorporating each $\bm x_i$ into the estimation of $\bm \Sigma_k$, we
increase the rank of $\widehat{\bm \Sigma}_k(\lambda)$ to $q \approx N$ if
$\lambda \ne 0$. Next, we provide an essential result that enables us to prove
that \eqref{eq:hdrda} is invariant to adjustments to the eigenvalues of $\tilde{
  \bm \Sigma}_k$ corresponding to the null space of $\widehat{\bm \Sigma}$.

\begin{lemma}\label{lemma:RDA-constant-term}
Let $\bm U_2$ be defined as above. Then, for all $\bm x \in \mathbb{R}_{p \times
  1}$, $\bm U_2^{T} (\bm x - \xbar_k) = \bm U_2^{T} (\bm x - \xbar_{k'})$, $1
\le k, k' \le K$, where $k \ne k'$.
\end{lemma}
\begin{proof}
Let $\bm x \in \mathbb{R}_{p \times 1}$, and suppose that $1 \le k, k' \le K$.
Recall that $\bm U_2 \in \mathcal{N}(\widehat{\bm \Sigma})$, which implies that
$\bm U_2^{T} \in \mathcal{C}(\widehat{\bm \Sigma})^{\perp}$ \citep[Lemma
  1.2.5]{Kollo:2005vp}. Now, because $\bm x_i \in \mathcal{C}(\widehat{ \bm
  \Sigma})$ $(i = 1, \ldots, N)$, $\bm U_2^{T} \bm x_i = \bm 0_{p-q}$.  Hence,
$\bm 0_{p-q} = \sum_{i=1}^N \beta_i \bm U_2^{T}\bm x_i = \bm U_2^{T}(\xbar_k -
\xbar_{k'})$, where $\beta_i = (n_k n_{k'})^{-1} \{ I(y_i = k) n_{k'} - I(y_i =
k') n_k \}$. Therefore, $\bm U_2^{T} (\bm x - \xbar_k) = \bm U_2^{T} (\bm x -
\xbar_{k'})$.
\end{proof}

We now present our main result, where we decompose \eqref{eq:hdrda} and show
that the term requiring the largest computational costs does not contribute to
the classification of an unlabeled observation performed using Lemma
\ref{lemma:RDA-constant-term}. Hence, we reduce \eqref{eq:hdrda} to an
equivalent, more computationally efficient decision rule.

\begin{thm}
Let $\tilde{\bm \Sigma}_k$ and $\bm W_k$ be defined as in \eqref{eq:rda-matrix}
and \eqref{eq:Wk}, respectively, and let $\bm U_1$ be defined as above. Then,
the decision rule in \eqref{eq:hdrda} is equivalent to
	\begin{align}
		D_{HDRDA}(\bm x) &= \argmin_k  (\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k) + \log | \bm W_k |. \label{eq:hdrda-decomposed}
	\end{align}
\end{thm}
\begin{proof}
From \eqref{eq:rda-matrix}, we have that $\tilde{\bm \Sigma}_k^{+} = \bm U(
\bm W_k^{-1} \oplus \gamma^{+} \bm I_{p-q} )\bm U^{T}$ and $|\tilde{\bm \Sigma}_k| =
\gamma^{p-q}| \bm W_k |$, $k = 1, \ldots, K$. Therefore, for all $\bm x \in
\mathbb{R}_{p \times 1}$, we have that
	\begin{align*}
	(\bm x - \xbar_k)^{T} \tilde{\bm \Sigma}_k^{+}(\bm x - \xbar_k)  + \log |\tilde{\bm \Sigma}_k| &= (\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k)\\
	&+ \gamma^{+} (\bm x - \xbar_k)^{T} \bm U_2 \bm U_2^{T} (\bm x - \xbar_k) + \log | \bm W_k |\\
  &+ (p - q) \log \gamma.
	\end{align*}
Because $\gamma$ is constant for $k = 1, \ldots, K$, we can omit the $(p - q)
\log \gamma$ term and particularly avoid the calculation of $\log 0$ for $\gamma
= 0$. Then, the proof follows from Lemma \ref{lemma:RDA-constant-term} because
$\bm U_2^{T} (\bm x - \xbar_k)$ is constant for $k = 1, \ldots, K$.
\end{proof}

Thus, we can avoid the time-consuming inverses and determinants of $p \times p$
covariance matrices in \eqref{eq:hdrda} and instead calculate these same
operations on $\bm W_k \in \mathbb{R}_{q \times q}$ in
\eqref{eq:hdrda-decomposed}. The substantial computational improvements arise
because our proposed classifier in \eqref{eq:hdrda} is invariant to the term
$\bm U_2$, thus yielding an equivalent classifier in \eqref{eq:hdrda-decomposed}
with a substantial reduction in computational complexity. Here, we demonstrate
that the computational efficiency in calculating the inverse and determinant of
$\bm W_k$ can be further improved via standard matrix operations when we show
that the inverses and determinants of $\bm W_k$ can be performed on matrices of
size $n_k \times n_k$.

\begin{proposition}\label{proposition:hdrda-W_k}
Let $\bm W_k$ be defined as above. Then, $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$
and
	\begin{align}
		\bm W_k^{-1} &= \bm \Gamma_k^{-1} - n_k^{-1} \alpha_k(1 - \lambda) \bm \Gamma_k^{-1} \bm U_1^{T} \bm X_k^{T} \bm Q_k^{-1} \bm X_k \bm U_1 \bm \Gamma_k^{-1},\label{eq:W_k_inv}
    \intertext{where}
    \bm Q_k &= \bm I_{n_k} + n_k^{-1} \alpha_k(1 - \lambda) \bm X_k \bm U_1 \bm \Gamma_k^{-1} \bm U_1^{T} \bm X_k^{T}\label{eq:Q_k}
    \intertext{and}
    \bm \Gamma_k &= \alpha_k \lambda \bm D_q + \gamma \bm I_q\label{eq:Gamma_k}.
	\end{align}
\end{proposition}
\begin{proof}
First, we write $\bm W_k = n_k^{-1} \alpha_k (1 - \lambda) \bm U_1^{T} \bm
X_k^{T} \bm X_k \bm U_1 + \bm \Gamma_k$. To calculate $|\bm W_k|$, we apply
Theorem 18.1.1 from \cite{Harville:2008wja}, which states that $|\bm A + \bm B
\bm T \bm C| = |\bm A| |\bm T| |\bm T^{-1} + \bm C \bm A^{-1} \bm B|$, where
$\bm A \in \mathbb{R}_{a \times a}^{>}$, $\bm B \in \mathbb{R}_{a \times b}$,
$\bm T \in \mathbb{R}_{b \times b}^{>}$, and $\bm C \in \mathbb{R}_{b \times
  a}$. Thus, setting $\bm A = \bm \Gamma_k$, $\bm B = \alpha_k (1 - \lambda) \bm
U_1^{T} \bm X_k^{T}$, $\bm T = \bm I_{n_k}$, and $\bm C = \bm X_k \bm U_1$, we
have $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$. Similarly, \eqref{eq:W_k_inv}
follows from the well-known Sherman-Woodbury formula \citep[Theorem
  18.2.8]{Harville:2008wja} by noting that $(\bm A + \bm B \bm T \bm C)^{-1} =
\bm A^{-1} - \bm A^{-1} \bm B (\bm T^{-1} + \bm C \bm A^{-1} \bm B)^{-1} \bm C
\bm A^{-1}$.
\end{proof}

The application of the Sherman-Woodbury formula requires that $\bm \Gamma_k$ be
nonsingular. This assumption is not met when $(\lambda, \gamma) = (0, 0)$
because $\bm \Gamma_k = \bm 0_q$. Here, we use the formulation in
\eqref{eq:hdrda-decomposed} instead. Also, notice that if $\alpha_k$ is constant
across the $K$ classes, then $\bm \Gamma_k$ in \eqref{eq:Gamma_k} is independent
of $k$. Consequently, $|\bm \Gamma_k|$ is constant across the $K$ classes and
need not be calculated in \eqref{eq:hdrda-decomposed}.

\subsection{Model Selection}

Thus far, we have presented the \emph{HDRDA} classifier and its properties that
facilitate an efficient calculation of the decision rule. Here, we describe an
efficient model-selection procedure along with pseudocode in Algorithm
\ref{alg:hdrda_model_selection} to select the optimal tuning-parameter estimates
from the Cartesian product of candidate values $\{\lambda_g\}_{g=1}^G \times
\{\gamma_h\}_{h=1}^H$. We estimate the $V$-fold cross-validation error rate for
each candidate pair and select $(\widehat{\lambda}, \widehat{\gamma})$, which
attains the minimum error rate. To calculate the $V$-fold cross-validation, we
partition the original training data into $V$ mutually exclusive and exhaustive
folds that have approximately the same number of observations. Then, for $v = 1,
\ldots, V$, we classify the observations in the $v$th fold by training a
classifier on the remaining $V - 1$ folds. We calculate the cross-validation
error as the proportion of misclassified observations across the $V$ folds.

\IncMargin{1em}
\begin{algorithm}
  \SetKwInOut{Input}{input}
  \SetKwInOut{Output}{output}

  \Input{
    Data matrix $\bm X$\\
    Parameter grid $\{\lambda_g\}_{g=1}^G \times \{\gamma_h\}_{h=1}^H$
  }
  \Output{Optimal Estimates $(\hat{\lambda}, \hat{\gamma})$}
  \BlankLine
  \For{$v \leftarrow 1$ \KwTo $V$} {
    Partition $\bm X$ into $\bm X_{train} \in \mathbb{R}_{N \times p}$ and
    $\bm X_{test} \in \mathbb{R}_{N_T \times p}$

    \For{$k\leftarrow 1$ \KwTo $K$}{
      Extract $\bm X_k \in \mathbb{R}_{n_k \times p}$ from $\bm X_{train}$

      Compute sample mean $\xbar_k$ from $\bm X_k$

      Center $\bm X_k \leftarrow \bm X_k - \bm 1_{n_k} \xbar_k^T$
    }

    $\bm X_c \leftarrow [\bm X_1^T, \ldots, \bm X_K^T]^T$

    Compute the compact SVD $\bm X_c = \bm M_q \bm D_q \bm U_1^T$

    Transform $\bm X_c \leftarrow \bm X_c \bm U_1$

    Transform $\bm X_{test} \leftarrow \bm X_{test} \bm U_1$

    \For{$k \leftarrow 1$ \KwTo $K$} {
      Extract $\bm X_k \in \mathbb{R}_{n_k \times q}$ from $\bm X_c$

      Recompute sample mean $\xbar_k$ from $\bm X_k$
    }

    \For{$(\lambda, \gamma) \in \{\lambda_g\}_{g=1}^G \times \{\gamma_h\}_{h=1}^H$} {
      \For{$k \leftarrow 1$ \KwTo $K$} {
        Compute $\bm Q_k$ using \eqref{eq:Q_k}

        Compute $\bm \Gamma_k$ using \eqref{eq:Gamma_k}

        Compute $\bm W_k^{-1}$ using \eqref{eq:W_k_inv}

        Compute $|\bm W_k| = |\bm \Gamma_k| |\bm Q_k|$

        Compute $(\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x -
        \xbar_k) + \log | \bm W_k |$ for each row $\bm x$ of $\bm X_{test}$
      }
      Classify test observations $\bm X_{test}$ using \eqref{eq:hdrda-decomposed}

      Compute the number of misclassified test observations $\#\{\text{Error}_v(\lambda, \gamma)\}$
    }
   }
  Compute $\widehat{\text{Error}}(\lambda, \gamma) = N^{-1} \sum_{v=1}^V \#\{\text{Error}_v(\lambda, \gamma)\}$

  Report optimal $(\hat{\lambda}, \hat{\gamma}) \leftarrow \argmin_{(\lambda, \gamma)} \widehat{\text{Error}}(\lambda, \gamma)$
  \caption{Model selection for the \emph{HDRDA} classifier}
  \label{alg:hdrda_model_selection}
\end{algorithm}
\DecMargin{1em}

A primary contributing factor to the efficiency of Algorithm
\ref{alg:hdrda_model_selection} is our usage of the compact singular value
decomposition (SVD). Rather than computing the eigenvalue decomposition of
$\widehat{\bm \Sigma}$ to obtain $\bm U_1$, we instead obtain $\bm U_1$ by
computing the eigendecomposition of a much smaller $N \times N$ matrix when $p
\gg N$ \cite[Chapter~18.3.5]{Hastie:2008dt}. Applying the SVD, we decompose $\bm
X_c = \bm M \bm \Delta \bm U^{T}$, where $\bm M \in \mathbb{R}_{N \times p}$ is
orthogonal, $\bm \Delta \in \mathbb{R}_{p \times p}^{\ge}$ is a diagonal matrix
consisting of the singular values of $\bm X_c$, and $\bm U \in \mathbb{R}_{p
  \times p}$ is orthogonal. Recalling that $\widehat{\bm \Sigma} = N^{-1} \bm
X_c^{T} \bm X_c$, we have the eigendecomposition $\widehat{\bm \Sigma} = \bm U
\bm D \bm U^{T}$, where $\bm U$ is the matrix of eigenvectors of $\widehat{\bm
  \Sigma}$ and $\bm D = N^{-1} \bm \Delta$ is the diagonal matrix of eigenvalues
of $\widehat{\bm \Sigma}$. Now, we can obtain $\bm M$ and $\bm D$ efficiently
from the eigenvalue decomposition of the $N \times N$ matrix $\bm X_c \bm
X_c^{T} = \bm M \bm D \bm M^{T}$. Next, we compute $\bm U = \bm X_c^{T} \bm M
\bm D^{+/2}$, where $\bm D^{+/2} = \bm D_q^{-1/2} \oplus \bm 0_{N-q}$. We then
determine $q$, the number of numerically nonzero eigenvalues present in $\bm D$,
by calculating the number of eigenvalues that exceeds some tolerance value, say,
$1 \times 10^{-6}$. We then extract $\bm U_1$ as the first $q$ columns of $\bm
U$.

As a result of the compact SVD, we need calculate $\bm X_c \bm U_1$ only once
per cross-validation fold, requiring $O(p q N) \approx O(p N^2)$ calculations.
Hence, the computational costs of expensive calculations, such as matrix
inverses and determinants, are greatly reduced because they are performed in the
$q$-dimensional subspace. Similarly, we reduce the dimension of the test data
set by calculating $\bm X_{test} \bm U_1$ once per fold. Conveniently, we see
that the most costly computation involved in $\bm Q_k$ and $\bm W_k^{-1}$ is
$\bm X_k \bm U_1$, which can be extracted from $\bm X_c \bm U_1$. Thus, after
the initial calculation of $\bm X_c \bm U_1$ per cross-validation fold, $\bm
Q_k$ requires $O(n_k q^2)$ operations. Because $\bm Q_k \in \mathbb{R}_{n_k
  \times n_k}$, both its determinant and inverse require $O(n_k^3)$
operations. Consequently, $\bm W_k^{-1}$ requires $O(n_k q^2)$ operations. Also,
the inverse of the diagonal matrix $\bm \Gamma_k^{-1} \in \mathbb{R}_{q \times
  q}$ requires $O(q)$ operations. Finally, we remark that $| \bm W_k |$ requires
$O(n_k^3)$ operations.

The expressions given in Proposition \ref{proposition:hdrda-W_k} also expedite
the selection of $\lambda$ and $\gamma$ via cross-validation because the most
time-consuming matrix operation involved in computing $\bm W_k^{-1}$ and $|\bm
W_k|$ is $\bm X_k \bm U_1 \in \mathbb{R}_{n_k \times q}$, which is independent
of $\lambda$ and $\gamma$. The subsequent operations in calculating $\bm
W_k^{-1}$ and $|\bm W_k|$ can be simply updated for different pairs of $\lambda$
and $\gamma$ without repeating the costly computations. When calculating the
discriminant scores of each test observation in $\bm X_{test}$, rather than
calculating each score individually, we can more efficiently calculate the
quadratic forms via a single matrix operation. That is, rather than calculating
$(\bm x - \xbar_k)^{T} \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm x - \xbar_k)$
individually for each row $\bm x$ of $\bm X_{test}$, we can calculate $(\bm
X_{test} - \xbar_k \bm 1_k')' \bm U_1 \bm W_k^{-1} \bm U_1^{T} (\bm X_{test} -
\xbar_k \bm 1_k')$. The diagonal elements of the resulting matrix contain the
individual quadratic form of each test observation, $\bm x_t$.

\section{Classification Study}
\label{sec:sims}

In this section, we compare our proposed classifier with four classifiers
recently proposed for small-sample, high-dimensional data along with the
random-forest classifier from \cite{Breiman:2001fb} using version 3.2.2 of the
open-source statistical software {\tt R}. Within our study, we included
penalized linear discriminant analysis from \cite{Witten:2011kc}, implemented in
the {\tt penalizedLDA} package. We also considered shrunken centroids
regularized discriminant analysis from \cite{Guo:2007te} in the {\tt rda}
package. Because the {\tt rda} package does not perform the authors' ``Min-Min''
rule automatically, we applied this rule within our {\tt R} code. We included
two modifications of diagonal linear discriminant analysis from
\cite{Tong:2012hw} and \cite{Pang:2009ik}, where the former employs an improved
mean estimator and the latter utilizes an improved variance estimator. Both
classifiers are available in the {\tt sparsediscrim} package. Finally, we
incorporated the random forest as a benchmark based on the findings of
\cite{FernandezDelgado:2014ul}, who concluded that the random forest is often
superior to other classifiers in benchmark studies. We used the implementation
of the random-forest classifier from the {\tt randomForest} package with 250
trees and 100 maximum nodes. For each classifer we explicitly set prior
probabilities as equal, if applicable. All other classifier options were set to
their default settings. Below, we refer to each classifier by the first author's
surname. All simulations were conducted on an Amazon Elastic Compute Cloud (EC2)
{\tt c4.4xlarge} instance.

For the \emph{HDRDA} classifier in \eqref{eq:hdrda-decomposed}, we examined the
classification performance of two models. For the first \emph{HDRDA} model, we
set $\alpha_k = 1$, $k = 1, \ldots, K$, so that the covariance-matrix estimator
\eqref{eq:hdrda-cov} resembled \eqref{eq:ridge-estimator}. We estimated
$\lambda$ from a grid of 21 equidistant candidate values between 0 and 1,
inclusively. Similarly, we estimated $\gamma$ from a grid consisting of the
values $10^{-1}, \ldots, 10^4$, and $10^5$. We selected optimal estimates of
$\lambda$ and $\gamma$ using $V$-fold cross-validation with $V = 10$
\citep{Hastie:2008dt}. For the second model, we set $\alpha_k = 1 - \gamma$, $k
= 1, \ldots, K$, to resemble \citeauthor{Friedman:1989tm}'s parameterization,
and we estimated both $\lambda$ and $\gamma$ from a grid of 21 equidistant
candidate values between 0 and 1, inclusively.

\subsection{Simulation Study}

In this section we compare the competing classifiers using the simulation design
from \cite{Guo:2007te}. This design is widely used within the high-dimensional
classification literature, including the studies by \cite{Ramey:2013ji} and
\cite{Witten:2011kc}. First, we consider the block-diagonal covariance matrix
from \cite{Guo:2007te},
\begin{align}
  \bm\Sigma_k = \bm\Sigma^{(\rho_k)} \oplus \bm\Sigma^{(-\rho_k)} \oplus \ldots \oplus \bm\Sigma^{(\rho_k)},\label{eq:block-sigma}
\end{align}
where the $(i,j)$th entry of the block matrix $\bm\Sigma^{(\rho_k)} \in
\mathbb{R}_{100 \times 100}$ is
\begin{align*}
\bm\Sigma_{ij}^{(\rho_k)} = \{ \rho_k^{|i - j|} \}_{1 \le i,j \le 100}.
\end{align*}
The block-diagonal covariance structure in \eqref{eq:block-sigma} resembles
gene-expression data: within each block of pathways, genes are correlated, and
the correlation decays as a function of the distance between any two genes. The
original design from \cite{Guo:2007te} comprised two $p$-dimensional
multivariate normal populations with a common block-diagonal covariance matrix
where $\bm\Sigma_1 = \bm\Sigma_2$.

Although the design is indeed standard, the simulation configuration lacks
artifacts observed commonly in real data, such as skewness and extreme
outliers. As a result, we wished to investigate the effect of outliers on the
high-dimensional classifiers. To accomplish this goal, we generalized the
block-diagonal simulation configuration by sampling from a $p$-dimensional
multivariate contaminated normal distribution.  Denoting the PDF of the
$p$-dimensional multivariate normal distribution by $N_p(\bm x | \bm \mu, \bm
\Sigma)$, we write the PDF of the $k$th class as
\begin{align}
  p(\bm x | \omega_k) = (1 - \epsilon) N_p(\bm x | \bm \mu_k, \bm \Sigma_k) + \epsilon N_p(\bm x | \bm \mu_k, \eta \bm \Sigma_k),\label{eq:contaminated-pdf}
\end{align}
where $\epsilon \in [0, 1]$ is the probability that an observation is
contaminated (i.e., drawn from a distribution with larger variance) and $\eta >
1$ scales the covariance matrix $\bm \Sigma_k$ to increase the extremity of
outliers. For $\epsilon = 0$, we have the benchmark block-diagonal simulation
design from \cite{Guo:2007te}. As $\epsilon$ is increased, the average number of
outliers is increased. In our simulation, we let $\eta = 100$ and considered the
values of $\epsilon = 0$, $0.05$, $\ldots$, $0.50$.

We generated $K=3$ populations from \eqref{eq:contaminated-pdf} with $\bm
\Sigma_k$ given in \eqref{eq:block-sigma} and set the mean vector of class 1 to
$\bm \mu_1 = \bm 0_p$. Next, comparable to \cite{Guo:2007te}, the first 100
features of $\bm \mu_2$ were set to 1/2, while the rest were set to 0, i.e.,
$\bm \mu_2 = (\underbrace{1/2, \ldots, 1/2}_{100}, \underbrace{0, \ldots, 0}_{p
  - 100})$. For simplicity, we defined $\bm \mu_3 = -\bm \mu_2$. The three
populations differed in their mean vectors in the first 100 features
corresponding to the first block. No difference in the means occurred in the
remaining blocks.

From each of the $K=3$ populations, we sampled 25 training observations ($n_k =
25$ for all $k$) and 10,000 test observations. After training each classifier on
the training data, we classified the test data sets and computed the proportion
of mislabeled test observations to estimate the classification error rate for
each classifier. Repeating this process 500 times, we computed the average of
the error-rate estimates for each classifier. We allowed the number of features
to vary from $p = 100$ to $p = 500$ in increments of 100 to examine the
classification accuracy as the feature dimension increased while maintaining a
small sample size. \cite{Guo:2007te} originally considered $\rho_k = 0.9$ for
all $k$. Alternatively, to explore the more realistic assumption of unequal
covariance matrices, we put $\rho_1 = 0.1$, $\rho_2 = 0.5$, and $\rho_3 = 0.9$

\subsubsection{Simulation Results}

In Figure \ref{fig:sim-results}, we observe each classifier's average
classification error rates for the values of $\epsilon$ and $p$. Unsurprisingly,
the average error rate increased for each classifier as the contamination
probability $\epsilon$ increased regardless of the value of $p$. Sensitivity to
the presence of outliers was most apparent for the Pang, Tong, and Witten
classifiers. For smaller dimensions, the random-forest and \emph{HDRDA}
classifiers tended to outperform the remaining classifiers with the random
forest performing best. As the feature dimension increased with $p \ge 300$,
both \emph{HDRDA} classifiers outperformed all other classifiers, suggesting
that their inherent dimension reduction better captured the classificatory
information in the small training samples, even in the presence of outliers.

\[ \left[\text{Insert Figure \ref{fig:sim-results} approximately here }\right] \]

The Pang, Tong, and Witten methods yielded practically the same and consistently
the worst error rates when outliers were present with $\epsilon > 0$, suggesting
that these classifiers were sensitive to outliers. Notice, for example, that
when $p=400$, the error rates of the Pang, Tong, and Witten classifiers
increased dramatically from approximately 19\% when no outliers were present to
approximately 43\% when $\epsilon=0.05$. The sharp increase in average error
rates for these three classifiers continued as $\epsilon$ increased. Guo's
method always outperformed those of Pang, Witten, and Tong, but after outliers
were introduced, the Guo classifier's average error rate was not competitive
with the \emph{HDRDA} classifiers or the random-forest classifier.

\[ \left[\text{Insert Figure \ref{fig:sim-results-by-p} approximately here }\right] \]

In Figure \ref{fig:sim-results-by-p} we again examine the simulation results as
a function of the number of features $p$ for a subset of the values of
$\epsilon$. This set of plots allows us to investigate the effect of feature
dimensionality on classification performance. When no outliers were present
(i.e., $\epsilon=0$), the random-forest classifier was outperformed by all other
classifiers. Furthermore, the \emph{HDRDA} classifiers were superior in terms of
average error rate in this setting. As $p$ increased, an elevation in average
error rate was expected for all classifiers, but the increase was not observed
to be substantial.

For $\epsilon > 0$, we observed a different behavior in classification
performance. First, the Pang, Tong, and Witten methods, along with the
random-forest method, increased in average error rate as $p$ increased.
Contrarily, the performance of the \emph{HDRDA} and Guo classifiers was hardly
affected by $p$. Also, as discussed above, the \emph{HDRDA} classifiers were
superior to all other classifiers for large values of $p$ with only the
random-forest classifier outperforming them in smaller feature-dimension cases.

\subsection{Application to Gene Expression Data}

We compared the \emph{HDRDA} classifier to the five competing classifiers on six
benchmark gene-expression microarray data sets. We evaluated the classification
accuracy of each classifier by randomly partitioning the data set under
consideration such that $2/3$ of the observations were allocated as training
data and the remaining $1/3$ of the observations were allocated as a test data
set. To expedite the computational runtime, we reduced the training data to the
top 1000 variables by employing the variable-selection method proposed by
\cite{Dudoit:2002ev}. We then reduced the test data set to the same 1000
variables. After training each classifier on the training data, we classified
the test data sets and computed the proportion of mislabeled test observations
to estimate the classification error rate for each classifier. Repeating this
process 100 times, we computed the average of the error-rate estimates for each
classifier. We next provide a concise description of each high-dimensional data
set examined in our classification study.

\subsubsection{\cite{Chiaretti:2004gq} Data Set}

\cite{Chiaretti:2004gq} measured the gene-expression profiles for 128
individuals with acute lymphoblastic leukemia (ALL) using Affymetrix human
95Av2 arrays. Following \cite{Xu:2009fl}, we restricted the data set to $K = 2$
classes such that $n_1 = 74$ observations were without cytogenetic abnormalities
and $n_2 = 37$ observations had a detected BCR/ABL gene. The robust multichip
average normalization method was applied to all 12,625 gene-expression levels.

\subsubsection{\cite{Chowdary:2006kf} Data Set}

\cite{Chowdary:2006kf} investigated 52 matched pairs of tissues from colon and
breast tumors using Affymetrix U133A arrays and ribonucleic-acid (RNA)
amplification. Each tissue pair was gathered from the same patient and consisted
of a snap-frozen tissue and a tissue suspended in an RNAlater preservative.
Overall, 31 breast-cancer and 21 colon-cancer pairs were gathered, resulting in
$K = 2$ classes with $n_1 = 62$ and $n_2 = 42$. A purpose of the study was to
determine whether the disease state could be identified using 22,283
gene-expression profiles.

\subsubsection{\cite{Nakayama:2007fl} Data Set}

\cite{Nakayama:2007fl} acquired 105 gene-expression samples of 10 types of
soft-tissue tumors through an oligonucleotide microarray, including 16 samples
of synovial sarcoma (SS), 19 samples of myxoid/round cell liposarcoma (MLS), 3
samples of lipoma, 3 samples of well-differentiated liposarcoma (WDLS), 15
samples of dedifferentiated liposarcoma (DDLS), 15 samples of myxofibrosarcoma
(MFS), 6 samples of leiomyosarcoma (LMS), 3 samples of malignant nerve sheathe
tumor (MPNST), 4 samples of fibrosarcoma (FS), and 21 samples of malignant
fibrous histiocytoma (MFH). \cite{Nakayama:2007fl} determined from their data
that these 10 types fell into 4 broader groups: (1) SS; (2) MLS; (3) Lipoma,
WDLS, and part of DDLS; (4) Spindle cell and pleomorophic sarcomas including
DDLS, MFS, LMS, MPNST, FS, and MFH. Following \cite{Witten:2011kc}, we restrict
our analysis to the five tumor types having at least 15 observations.

\subsubsection{\cite{Shipp:2002ka} Data Set}

According to \cite{Shipp:2002ka}, approximately 30\%-40\% of adult non-Hodgkin
lymphomas are diffuse large B-cell lymphomas (DLBCLs). However, only a small
proportion of DLBCL patients are cured with modern chemotherapeutic regimens.
Several models have been proposed, such as the International Prognostic Index
(IPI), to determine a patient's curability.  These models rely on clinical
covariates, such as age, to determine if the patient can be cured, and the
models are often ineffective. \cite{Shipp:2002ka} have argued that researchers
need more effective means to determine a patient's curability. The authors
measured 6,817 gene-expression levels from 58 DLBCL patient samples with
customized cDNA (lymphochip) microarrays to investigate the curability of
patients treated with cyclophosphamide, adriamycin, vincristine, and prednisone
(CHOP)-based chemotherapy. Among the 58 DLBCL patient samples, 32 are from cured
patients while 26 are from patients with fatal or refractory disease.

\subsubsection{\cite{Singh:2002fh} Data Set}

\cite{Singh:2002fh} have examined 235 radical prostatectomy specimens from
surgery patients between 1995 and 1997. The authors used oligonucleotide
microarrays containing probes for approximately 12,600 genes and expressed
sequence tags. They have reported that 102 of the radical prostatectomy
specimens are of high quality: 52 prostate tumor samples and 50 non-tumor
prostate samples.

\subsubsection{\cite{Tian:2003ht} Data Set}

\cite{Tian:2003ht} investigated the purified plasma cells from the bone marrow
of control patients along with patients with newly diagnosed multiple
myeloma. Expression profiles for 12,2625 genes were obtained via Affymetrix
U95Av2 microarrays. The plasma cells were subjected to biochemical and
immunohistochemical analyses to identify molecular determinants of osteolytic
lesions. For 36 multiple-myloma patients, focal bone lesions could not be
detected by magnetic resonance imaging (MRI), whereas MRI was used to detect
such lesions in 137 patients.

\subsubsection{Classification Results}

Similar to \cite{Witten:2011kc}, we report the average test error rates obtained
over 100 random training-test partitions in Table \ref{tab:microarray-results}
along with standard deviations of the test error rates in parentheses. The
\emph{HDRDA} and Guo classifiers were superior in classification performance for
the majority of the simulations. The \emph{HDRDA} classifiers had the best
classification accuracy on the Chowdary and Shipp data sets. Although the random
forest's accuracy slightly exceeded the \emph{HDRDA} classifiers on the Tian
data set, our proposed classifiers outperformed the other competing classifiers
considered here. Moreover, the \emph{HDRDA} classifiers had comparable
performance on five of the six data sets.

\[ \left[\text{Insert Table \ref{tab:microarray-results} approximately here }\right] \]

The average error-rate estimates for the Pang, Tong, and Witten classifiers were
comparable across all six data sets. Furthermore, the average error rates for
the Pang and Tong classifiers were approximately equal for all data sets except
for the Chiaretti dataset. This result suggests that the mean and variance
estimators used in lieu of the MLEs provided little improvement to
classification accuracies. However, we investigated the Pang classifier's poor
performance on the Chiaretti data set and determined that its variance estimator
exhibited numerical instability. The classifier's denominator was approximately
zero for both classes and led to the poor classification performance.

The random-forest classifier was competitive when applied to the Chowdary and
Singh data sets and yielded the smallest error rate of the considered
classifiers on the Tian data set. The fact that the \emph{HDRDA} and Guo
classifiers typically outperformed the random-forest classifier challenges the
claim of \cite{FernandezDelgado:2014ul} that random forests are typically
superior. Further studies should be performed to validate this statement in the
small-sample, high-dimensional setting.

Finally, the Pang, Tong, and Witten classifiers consistently yielded the largest
average error rates across the six data sets. Given that the standard deviations
were relatively large, we hesitate to generalize claims regarding the ranking of
these three classifiers in terms of the average error rate. However, the
classifiers' error rates and their variability across multiple random partitions
of each data set were large enough that we might question their benefit when
applied to real data.

\section{Discussion}
\label{sec:discussion}

We have compared our proposed \emph{HDRDA} classifier to the random-forest and
Witten classifiers as well as the Pang, Tong, and Guo classifiers, each of which
assumes independence of the features conditional on class membership. Despite
the rapid computational performance of diagonal classifiers and the reduction in
the number of parameters to estimate, \cite{Mai:2012bf} and \cite{Fan:2012iq}
have noted that diagonal classifiers can often yield inferior classification
performance when compared to other classification methods. In fact, we have
shown that the \emph{HDRDA} classifier often yields superior classification
accuracy when compared to the diagonal classifiers considered using six
small-sample, high-dimensional data sets, confirming the assertions of
\cite{Mai:2012bf} and \cite{Fan:2012iq}. Furthermore, we have demonstrated that
\emph{HDRDA} classifiers are more robust to the presence of outliers than the
diagonal classifiers.

We also considered the popular penalized linear discriminant analysis from
\cite{Witten:2011kc} because it was specifically designed for high-dimensional
gene-expression data. We had expected its classification performance to be
competitive within our classification study and perhaps superior. Contrarily,
its underperformance in our empirical studies show that it is sensitive to
outliers and unable to achieve comparable results with other classifiers
designed for small-sample, high-dimensional data. Also, despite the claims of
\cite{FernandezDelgado:2014ul} that the random forest classifier is typically
superior to other classifiers, we observed that the random-forest classifier was
indeed competitive but was typically outperformed by classifiers developed for
small-sample, high-dimensional data.

The \emph{HDRDA} implementation of \eqref{eq:hdrda-decomposed} in the {\tt
  sparsediscrim} R package was comparable to that of the diagonal classifiers in
terms of computational runtime. Given the \emph{RDA} classifier from
\cite{Friedman:1989tm} has been shown to have excellent performance in the
high-dimensional setting \citep{Webb:2011vu} but is limited by its
computationally intense model-selection procedure, our work enables the
\emph{RDA} classifier to be employed with small-sample, high-dimensional data in
practice. This result is reassuring because the \emph{RDA} classifier remains
widely popular in the literature. In fact, variants of the RDA classifier have
been applied to microarray data
\citep{Ching:2012fu,Li:2012ev,Tai:2007bk,Guo:2007te}, facial recognition
\citep{Ching:2012fu,Zhang:2010va,Dai:2007vd,Lu:2005hq,Pima:2004vw,Lu:2003we},
handwritten digit recognition \citep{Bouveyron:2007gx}, remote sensing
\citep{Tadjudin:1999fk}, seismic detection \citep{Anderson:2002kg}, and chemical
spectra \citep*{Wu:1996us,Aeberhard:1993fp}.

The dimension reduction that we have employed in this paper has reduced the
dimension to rank$(\widehat{\bm \Sigma}) = q$. An interesting extension of our
work would reduce the dimension $q$ further to a lower dimension $q_L < q$ using
some criterion proposed, perhaps similar to that employed in principal
components analysis. While unclear whether the classification performance would
improve via such a method, the efficiency of the training and model selection
would certainly improve. Moreover, if $q_L = 2$ or $3$ was selected, the
potential for low-dimensional graphical displays of high-dimensional data could
be accomplished.

\bibliographystyle{plainnat}
\bibliography{rda}

\clearpage

\begin{figure}
<<hdrda_contours, fig.width=8, fig.height=8, out.width='.33\\linewidth', dpi=200, fig.show='hold'>>=
rda_contours(lambda = 0)
rda_contours(lambda = 0.2)
rda_contours(lambda = 0.4)
rda_contours(lambda = 0.6)
rda_contours(lambda = 0.8)
rda_contours(lambda = 1)
@
\caption{Contours of five multivariate normal populations as a function of the
  pooling parameter $\lambda$.}
\label{fig:hdrda-contours}
\end{figure}

\clearpage

\begin{figure}
<<sim_results, fig.width=8, fig.height=8, dpi=200>>=
p <- ggplot(results_contaminated, aes(x=contamination_prob, y=Error))
p <- p + geom_line(aes(color=Classifier, linetype=Classifier), size=0.65)
p <- p + facet_wrap(~ p, labeller = "label_both")
p <- p + xlab("Contamination Probability") + ylab("Average Error Rate")
p + theme_bw()
@
\caption{Average classification error rates as a function of the contamination
  probability $\epsilon$. Approximate standard errors were no greater than
  \Sexpr{max_sim_se}.}
\label{fig:sim-results}
\end{figure}

\clearpage

\begin{figure}
<<sim_results_by_p, fig.width=8, fig.height=8, dpi=200>>=
p2 <- results_contaminated %>%
  mutate(p=as.numeric(as.character(p)),
         contamination_prob=round(contamination_prob, digits=2)) %>%
    filter(contamination_prob %in% c(0.0, 0.1, 0.2, 0.3, 0.4, 0.5)) %>%
    rename(`Contamination Probability`=contamination_prob) %>%
      ggplot(., aes(x=p, y=Error))
p2 <- p2 + geom_line(aes(color=Classifier, linetype=Classifier), size=0.65)
p2 <- p2 + facet_wrap(~ `Contamination Probability`, labeller="label_both")
p2 <- p2 + xlab("Number of Features") + ylab("Average Error Rate")
p2 + theme_bw()
@
\caption{Average classification error rates as a function of the number of
  features $p$. Approximate standard errors were no greater than
  \Sexpr{max_sim_se}.}
\label{fig:sim-results-by-p}
\end{figure}

\clearpage

<<classification_results, results='asis'>>=
table_caption <- 'The average of the test error rates obtained on gene-expression data sets over 100 random training-test partitions. Standard deviations of the test error rates are given in the parentheses. The classifier with the minimum average error rate for each data set is in bold.'

xtable_out <- xtable(results_microarray,
                     caption=table_caption,
                     label='tab:microarray-results')
print(xtable_out, include.rownames=FALSE, sanitize.text.function = function(x) { x })
@

\end{document}