### 2017-02-05 reports/title-page.tex
Added title page for Pattern Recognition.

\documentclass[review,authoryear]{elsarticle}
\usepackage{url}

%----------------Page dimensions ----------------
  \oddsidemargin 0.0in
\evensidemargin 0.0in
\topmargin -0.75in
\leftmargin 0in
\headheight 0.0in
\headsep 0.5in
\footskip 0.5in
\footnotesep 0.0in
\textwidth 6.5in
\textheight 9.5in

\begin{document}

\begin{frontmatter}

\title{High-Dimensional Regularized Discriminant Analysis}

\author[jar]{John A. Ramey}
\author[cks]{Caleb K. Stein}
\author[pdy]{Phil D. Young}
\author[dmy]{Dean M. Young}
\address[jar]{\textbf{Corresponding Author} \\ Novi Labs \\ 7020 Easy Wind Drive, Suite 210 \\ Austin, TX 78752 \\ USA \\ Email: johnramey@gmail.com}
\address[cks]{Myeloma Institute \\ University of Arkansas for Medical Sciences \\ 4018 W. Capitol Avenue \\ Little Rock, AR 72205 \\ USA \\ Email: stein.caleb@gmail.com}
\address[pdy]{Department of Management and Information Systems \\ Baylor University \\ One Bear Place \#98005 \\ Waco, TX 76798-8005 \\ USA \\ Email: Philip\_Young@baylor.edu}
\address[dmy]{Department of Statistical Science \\ Baylor University \\ P.O. Box 97140 \\ Waco, TX 76798-7140 \\ USA \\ Email: Dean\_Young@baylor.edu}

\end{frontmatter}

\end{document}