### 2017-02-12 reports/highlights.tex
Added paper's highlights for Pattern Recognition

\documentclass{elsarticle}

%----------------Page dimensions ----------------
  \oddsidemargin 0.0in
\evensidemargin 0.0in
\topmargin -0.75in
\leftmargin 0in
\headheight 0.0in
\headsep 0.5in
\footskip 0.5in
\footnotesep 0.0in
\textwidth 6.5in
\textheight 9.5in

\begin{document}

\title{Highlights}

\maketitle

\begin{itemize}
  \item We present an efficient classifier designed for small-sample, high-dimensional data.
  \item Interpretable alternative to Friedman's Regularized Discriminant Analysis.
  \item HDRDA's matrix operations are linear in the number of features.
  \item Timing comparisons show our classifier is substantially faster than Friedman's RDA.
  \item Often superior to sparse and regularized classifiers in classification accuracy.
\end{itemize}

\end{document}