### 2017-10-19 reports/title-page.tex
updated title page

\documentclass[]{interact}
\usepackage{url}

\begin{document}

\articletype{Title Page}
\title{High-Dimensional Regularized Discriminant Analysis}

\author{
\name{John A. Ramey\textsuperscript{a}\thanks{CONTACT John A. Ramey. Email: johnramey@gmail.com}, Caleb K. Stein\textsuperscript{b}, Phil D. Young\textsuperscript{c}, and Dean M. Young\textsuperscript{d}}
\affil{\textsuperscript{a}Novi Labs, Austin, TX, USA;\\ \textsuperscript{b}Myeloma Institute, University of Arkansas for Medical Sciences, Little Rock, AR, USA;\\ \textsuperscript{c}Department of Management and Information Systems, Baylor University, Waco, TX, USA;\\ \textsuperscript{d}Department of Statistical Science, Baylor University, Waco, TX, USA}
}

\maketitle

\end{document}